### 2011-10-31 Setup_honeypot.tex
Opzet Setup Document

% !TEX TS-program = pdflatex
% !TEX encoding = UTF-8 Unicode

% This is a simple template for a LaTeX document using the "article" class.
% See "book", "report", "letter" for other types of document.

\documentclass[11pt]{article} % use larger type; default would be 10pt

\usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)

%%% Examples of Article customizations
% These packages are optional, depending whether you want the features they provide.
% See the LaTeX Companion or other references for full information.

%%% PAGE DIMENSIONS
\usepackage{geometry} % to change the page dimensions
\geometry{a4paper} % or letterpaper (US) or a5paper or....
% \geometry{margin=2in} % for example, change the margins to 2 inches all round
% \geometry{landscape} % set up the page for landscape
%   read geometry.pdf for detailed page layout information

\usepackage{graphicx} % support the \includegraphics command and options

% \usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent

%%% PACKAGES
\usepackage{booktabs} % for much better looking tables
\usepackage{array} % for better arrays (eg matrices) in maths
\usepackage{paralist} % very flexible & customisable lists (eg. enumerate/itemize, etc.)
\usepackage{verbatim} % adds environment for commenting out blocks of text & for better verbatim
\usepackage{subfig} % make it possible to include more than one captioned figure/table in a single float
% These packages are all incorporated in the memoir class to one degree or another...

%%% HEADERS & FOOTERS
\usepackage{fancyhdr} % This should be set AFTER setting up the page geometry
\pagestyle{fancy} % options: empty , plain , fancy
\renewcommand{\headrulewidth}{0pt} % customise the layout...
\lhead{}\chead{}\rhead{}
\lfoot{}\cfoot{\thepage}\rfoot{}

%%% SECTION TITLE APPEARANCE
\usepackage{sectsty}
\allsectionsfont{\sffamily\mdseries\upshape} % (See the fntguide.pdf for font help)
% (This matches ConTeXt defaults)

%%% ToC (table of contents) APPEARANCE
\usepackage[nottoc,notlof,notlot]{tocbibind} % Put the bibliography in the ToC
\usepackage[titles,subfigure]{tocloft} % Alter the style of the Table of Contents
\renewcommand{\cftsecfont}{\rmfamily\mdseries\upshape}
\renewcommand{\cftsecpagefont}{\rmfamily\mdseries\upshape} % No bold!

%%% END Article customizations

%%% The "real" document content comes below...

\title{Setup Honeypot: Team Contagiarium}
\author{Mark Vijfvinkel, Aram Verstegen, Joost Kremers \& Bart Lutgens}
%\date{} % Activate to display a given date or no date (if empty),
         % otherwise the current date is printed

\begin{document}
\maketitle

\section{Introduction}

In this document we explain how we have setup the utility machine and the honeypot. 
We have decided to split up our monitoring time into two parts. In the first part we configure the honeypot as a trapped based system and monitor it for 1 month.
Then we will reconfigure the honeypot to a completely open system and moniter that for 1 month. 
After that period we will analyze the results of both setups and compare the differences in the results we have captured. 
The advantage of a trap based system is that the honeypot can only be compromised to a certain extent and not really used for any malicious activity.
The disadvantage is that it is not possible to capture all the malicious activity an attacker can perform.
The advantage of a completely open system is that you can capture all the malicious activity, but that the system might also be completely compromised and do damage to other systems.

\section{Setup: Utility Machine}

We have configured this machine to log syslog messages coming from the honeypot.

\subsection{Setup: Honeypot}

More text.

\end{document}

### 2011-10-31 Setup_honeypot.tex
Heb het eea toegevoegd

@@
 \usepackage{subfig} % make it possible to include more than one captioned figure/table in a single float
+\usepackage{appendix}
 % These packages are all incorporated in the memoir class to one degree or another...
@@
 
-We have configured this machine to log syslog messages coming from the honeypot.
+In this section we decribe the setup of the utility machine.
+
+\subsection{Syslog}
+Syslog uses UDP to transmit its logfiles to a remote server and this is not very desirable, because everything is sent in plaintext and delivery is not garanteed.
+A VPN session between the two hosts is possible, but might be considered as overkill and still does not garantee delivery.
+Therefore we have configured syslog to use TCP, which makes it possible to use SSH to tunnel the TCP traffic.
+We are using a reverse tunnel from the utility machine to the honeypot.
+We use a script to allow the connection to be re-esthablished quickly when the SSH session is terminated for some reason. 
+\cite{Remote_logging}
 
@@
 
+\bibliographystyle{alpha}
+\bibliography{ref_bib}
+
 \end{document}

### 2011-11-01 Setup_honeypot.tex
extended some text

@@
 
-\subsection{Setup: Honeypot}
+\subsection{Root}
 
-More text.
+\section{Setup: Honeypot}
+
+In this section we decribe the setup of the honeypot.
+
+\subsection{scanlogd}
+
+To detect portscans we have installed the tool scanlogd.
+Per scan an entry is written to syslog.
+
+\subsection{iptables}
+To log all connection attempts coming in and going out, we have added the following rules to the iptables configuration. 
+This will also log to syslog.
+
+\emph{iptables -I INPUT -m state --state NEW -j LOG --log-prefix "New Connection: "}
+
+\emph{iptables -I OUTPUT -m state --state NEW -j LOG --log-prefix "New Connection: "}
+
+\subsection{Kippo}
+To capture attackers that try to come in over ssh, we are using Kippo. 
+Kippo is used to log bruteforce attacks and the shell interaction performed by the attacker. 
+To make this as realistic as possible we have added a redicrect in the iptables configuration from port 22 to port 2222 that kippo uses. 
+Unfortunately this broke down after a while, for reasons yet unknown. 
+The reason for forwarding was not to give Kippo root capabilities.
+
+To be able to still log on to the honeypot, we configured port 9022 to be used by ssh.
+
+\subsection{Google Hack Honeypot}
+Google Hack Honeypot (GHH) is used to attract search engine attackers. 
+These attackers will use search queries to find sensitive data, for example password files, that due to some misconfiguration ended up being indexed by a search engine like google.
+To set this up we have first installed Apache and PHP5.
+Then we configured GHH to use Comma Seperated Values, since MySQL was not installed yet on the util machine and to see if this would actually work. 
+We can later reconfigure to use MySQL, if the results are not satisfactory.
+Since GHH offers multiple setup honeypot we picked one to start with, namely GGH version 1.2 passlist.txt.
+We had no particular reason for this, only that it was easy to set up. 
+The plan was to install more variants of the GHH to attract more attackers, but before this could happen the VM's were shut down.
+
+\subsection{Snort}
+
+Stukje Joost en Bart.
 

### 2011-11-01 Setup_honeypot.tex
snort

@@
 \subsection{Snort}
-
-Stukje Joost en Bart.
+Snort is a network intrusion detection system (NIDS) that uses realtime traffic analysis and packet logging on internet protocol networks.
+we use it to detect several attacks including stealth port scans, buffer overflows and more.
+Snort can be used as a network intrusion prevention system, however we will only be using it to detect attacks.
+Snort is rule-based, but we only intend to setup a few general rules that will only alert to suspicious network behaviour and do nothing to prevent it. 
+If a rule is triggered, we can look at the logs from the packet logger to analyse the network traffic.
+Snort sends alerts and packet logging files to syslog and as stated earlier, the syslog log files are transfered to the remote util machine.
 

### 2011-11-01 Setup_honeypot.tex
snort2

@@
 Snort can be used as a network intrusion prevention system, however we will only be using it to detect attacks.
-Snort is rule-based, but we only intend to setup a few general rules that will only alert to suspicious network behaviour and do nothing to prevent it. 
+Snort is rule-based and we intend to setup snort using a predefined rules-set that will only alert to suspicious network behaviour and do nothing to prevent it. 
 If a rule is triggered, we can look at the logs from the packet logger to analyse the network traffic.

### 2011-11-01 Setup_honeypot.tex
Update Snort.

--- 2011-11-01 Setup_honeypot.tex
@@
 Snort is a network intrusion detection system (NIDS) that uses realtime traffic analysis and packet logging on internet protocol networks.
-we use it to detect several attacks including stealth port scans, buffer overflows and more.
 Snort can be used as a network intrusion prevention system, however we will only be using it to detect attacks.
-Snort is rule-based, but we only intend to setup a few general rules that will only alert to suspicious network behaviour and do nothing to prevent it. 
+Snort is rule-based and we intend to setup Snort using a predefined rules-set that will only alert on suspicious network behaviour and do nothing to prevent it. 
+The rules that we will be using are based on a honeypot rule-set.
+Included are rules on detecting backdoors, viruses, DDOS, spam, SMTP- and other exploits.
+There rules will be tweaked if the results are poor.
 If a rule is triggered, we can look at the logs from the packet logger to analyse the network traffic.

### 2011-11-01 Setup_honeypot.tex
Kleine foutjes.

@@
 We have decided to split up our monitoring time into two parts. In the first part we configure the honeypot as a trapped based system and monitor it for 1 month.
-Then we will reconfigure the honeypot to a completely open system and moniter that for 1 month. 
+Then we will reconfigure the honeypot to a completely open system and monitor that for 1 month. 
 After that period we will analyze the results of both setups and compare the differences in the results we have captured. 
-The advantage of a trap based system is that the honeypot can only be compromised to a certain extent and not really used for any malicious activity.
+The advantage of a trap based system is that the honeypot can only be compromised to a certain extent and can not really be used for any malicious activity.
 The disadvantage is that it is not possible to capture all the malicious activity an attacker can perform.

### 2011-11-01 Setup_honeypot.tex
Subsection DenyHosts toegevoegd. Misschien nog wat uitbereiding nodig?

@@
 
+\subsection{DenyHosts}
+If someone finds out our real SSH port changed to 9022, they might try to break into this port.
+To make sure no bruteforce attacks can be done on this port, DenyHosts is installed.
+This program blocks IP-adresses that try to log in more than 3 times, with the wrong password.
+The blocked hosts are blacklisted on the honeypot, so we can edit them if needed.
+
 \subsection{Google Hack Honeypot}

### 2011-11-01 Setup_honeypot.tex
Reworded some things, added some things, fixed spelling mistakes

--- 2011-11-01 Setup_honeypot.tex
@@
 
-In this document we explain how we have setup the utility machine and the honeypot. 
-We have decided to split up our monitoring time into two parts. In the first part we configure the honeypot as a trapped based system and monitor it for 1 month.
-Then we will reconfigure the honeypot to a completely open system and moniter that for 1 month. 
-After that period we will analyze the results of both setups and compare the differences in the results we have captured. 
-The advantage of a trap based system is that the honeypot can only be compromised to a certain extent and not really used for any malicious activity.
-The disadvantage is that it is not possible to capture all the malicious activity an attacker can perform.
-The advantage of a completely open system is that you can capture all the malicious activity, but that the system might also be completely compromised and do damage to other systems.
+In this document we explain how we have set up the our honeypot. 
+We have decided to split up our monitoring time into two parts.
+In the first part we configure the virtual machine as a system of security `traps', also called a low-interaction (LI) honeypot, and monitor activity for one month.
+Then we will reconfigure the machine to a high-interaction (HI) honeypot, exposing real, exploitable vulnerabilities that can lead to code execution on our machine and monitor much more intently for another month.
 
-\section{Setup: Utility Machine}
+After these period we will analyze the results of both setups and compare the differences in the results we have captured. 
+The advantage of a LI honeypot is that the machine can only be compromised to a certain extent and can not really be used for any malicious activity.
+The disadvantage is that it is not possible to capture all the activity an attacker might perform, like after escalating privileges - which should not be possible.
+The advantage of a HI honeypot is that you can capture all the activity, but that the system might also be completely compromised and do damage to other systems.
+That is why we need to be much more vigilant in monitoring our honeypot in the HI phase.
 
-In this section we decribe the setup of the utility machine.
+\section{Setup: Monitoring Machine}
+
+In this section we describe the setup of the monitoring machine.
 
 \subsection{Syslog}
-Syslog uses UDP to transmit its logfiles to a remote server and this is not very desirable, because everything is sent in plaintext and delivery is not garanteed.
-A VPN session between the two hosts is possible, but might be considered as overkill and still does not garantee delivery.
-Therefore we have configured syslog to use TCP, which makes it possible to use SSH to tunnel the TCP traffic.
-We are using a reverse tunnel from the utility machine to the honeypot.
-We use a script to allow the connection to be re-esthablished quickly when the SSH session is terminated for some reason. 
-\cite{Remote_logging}
+By default, \verb|syslog| uses UDP to transmit its log messages to a remote server.
+Relying on UDP is not very desirable, because everything is sent in plaintext and delivery of UDP messages is not guaranteed.
+To allow for safe and reliable transmission of the syslog messages from the honeypot to the monitoring machine, one needs to use TCP connections in syslog, and provide some way of encapsulating the socket used in a cryptographic layer.
 
-\subsection{Root}
+%A VPN session between the two hosts is possible, but might be considered as overkill and still does not garantee delivery.
+We have decided to set up a reverse SSH tunnel from the monitoring machine to the honeypot machine, as described in the article ``Remote logging with syslog-ng''. \cite{Remote_logging}
+The command establishes an SSH connection the honeypot, setting options to forego a login shell or anything a normal user would desire, and binds the local socket 9514 on the honeypot machine to tunnel its connections to the same port on our monitoring machine.
+We have put this command there in a script to allow the connection to be re-esthablished quickly and easily if the SSH session is terminated for some reason.
+
+We have configured a locally installed the \verb|syslog-ng| package for our \verb|util| account user, configured it to listen to local TCP connections on port 9514 and thereby receive syslog messages transmitted through this port.
+
+To allow for secure SSH access, a public keypair was generated on the util machine and the associated public key was added to the honeypot's \verb|authorized_keys| file.
+
+ %which makes it possible to use SSH to tunnel the TCP traffic.
+
+%\subsection{Root}
+% haha surely you jest :>
 
@@
 
+\subsection{Rsyslog}
+On the honeypot, we've modified the \verb|rsyslog| configuration to also send syslog messages to the (now bound) local port 9514, completing the chain.
+The modified configuration is now stored in its own file in the syslog.d directory, which is a bit obvious to anyone taking interest.
+The configuration should and will be moved over to one of the default configuration files to avoid raising suspicion.
+Note that the chosen port is very similar to the default syslog port 514, and this might give us away.
+We plan to change it to a more nondescript port number.
+
 \subsection{scanlogd}
-
-To detect portscans we have installed the tool scanlogd.
-Per scan an entry is written to syslog.
+To detect port scans we have installed the service scanlogd.
+Per detected scan an entry is written to syslog.
 
 \subsection{iptables}
-To log all connection attempts coming in and going out, we have added the following rules to the iptables configuration. 
+To log all incoming and outgoing connection attempts, we have added the following rules to the iptables configuration. 
 This will also log to syslog.
@@
 \emph{iptables -I INPUT -m state --state NEW -j LOG --log-prefix "New Connection: "}
-
 \emph{iptables -I OUTPUT -m state --state NEW -j LOG --log-prefix "New Connection: "}
@@
 \subsection{Kippo}
-To capture attackers that try to come in over ssh, we are using Kippo. 
+To capture attackers that try to come in by brute-forcing SSH logins, we are using Kippo. 
 Kippo is used to log bruteforce attacks and the shell interaction performed by the attacker. 
-To make this as realistic as possible we have added a redicrect in the iptables configuration from port 22 to port 2222 that kippo uses. 
-Unfortunately this broke down after a while, for reasons yet unknown. 
+To make this as realistic as possible we have added a redirect in the iptables configuration from port 22 to port 2222 that kippo uses. 
+Unfortunately this broke down around the time the firewall was being reconfigured, for unclear reasons. 
 The reason for forwarding was not to give Kippo root capabilities.
 
-To be able to still log on to the honeypot, we configured port 9022 to be used by ssh.
+To be able to still log on to the honeypot, we configured SSH to use port 9022.
+
+\subsection{DenyHosts}
+If someone finds out our real SSH port changed to 9022, they might try to break in via our real SSH service.
+To make sure no bruteforce attacks can be done on this port, DenyHosts is installed.
+This service blocks IP-adresses that try to log in more than 3 times with the wrong password by adding them to the \verb|hosts.deny| file.
+The hosts listed there are blacklisted on the machine, but we can edit them if needed.
 
 \subsection{Google Hack Honeypot}
-Google Hack Honeypot (GHH) is used to attract search engine attackers. 
-These attackers will use search queries to find sensitive data, for example password files, that due to some misconfiguration ended up being indexed by a search engine like google.
-To set this up we have first installed Apache and PHP5.
-Then we configured GHH to use Comma Seperated Values, since MySQL was not installed yet on the util machine and to see if this would actually work. 
-We can later reconfigure to use MySQL, if the results are not satisfactory.
-Since GHH offers multiple setup honeypot we picked one to start with, namely GGH version 1.2 passlist.txt.
+Google Hack Honeypot (GHH) is designed to be used to attract search engine attackers. 
+These attackers will use search queries to find sensitive data, for example password files, that due to some misconfiguration ended up being indexed by a search engine like Google.
+GHH allows logging of these requests to a MySQL database or Comma Seperated Value (CSV) file.
+
+Since GHH offers multiple search targets, we picked one to start with, namely GGH version 1.2 passlist.txt.
 We had no particular reason for this, only that it was easy to set up. 
-The plan was to install more variants of the GHH to attract more attackers, but before this could happen the VM's were shut down.
+We plan to install more variants of the GHH to attract more attackers, but before this could happen the VM's were shut down.
+
+To set up GHH we have first installed Apache and PHP5.
+We configured GHH to log to in CSV format, since it requires minimal setup and is easily parsed.
+%We can later reconfigure to use MySQL, if the results are not satisfactory.
+Then, we patched the GHH code to allow logging to syslog, to maintain a single, consistent point of log collection.
 
 \subsection{Snort}
-Snort is a network intrusion detection system (NIDS) that uses realtime traffic analysis and packet logging on internet protocol networks.
-Snort can be used as a network intrusion prevention system, however we will only be using it to detect attacks.
-Snort is rule-based and we intend to setup Snort using a predefined rules-set that will only alert on suspicious network behaviour and do nothing to prevent it. 
-The rules that we will be using are based on a honeypot rule-set.
-Included are rules on detecting backdoors, viruses, DDOS, spam, SMTP- and other exploits.
-There rules will be tweaked if the results are poor.
-If a rule is triggered, we can look at the logs from the packet logger to analyse the network traffic.
-Snort sends alerts and packet logging files to syslog and as stated earlier, the syslog log files are transfered to the remote util machine.
+Snort is an intrusion detection system (IDS) that uses realtime traffic analysis and packet logging on internet protocol networks.
+Snort was designed to be used as a network IDS, however we will be running it on our host machine, effectively making it a host-based IDS.
+
+Unfortunately the VM network went down before we were able to complete our Snort configuration.
+We intend to set up Snort using a predefined rule-set that will only alert on suspicious network behaviour, but do nothing to prevent it. 
+The rules that we will be using are based on a honeypot rule-set. % TODO reference
+Included are rules on detecting backdoors, viruses, DDOS, spam and other abuse.
+These rules will be tweaked if the results are unsatisfactory.
+If a rule is triggered, snort will log an alert.
+
+We will configure Snort to send alerts to syslog, and to log its packet captures to some form of read-only filesystem mounted from our monitoring machine (yet to be decided).
+We will install a script like \verb|auto-oink| to automatically retrieve the latest community Snort rules.
 

### 2011-11-01 Setup_honeypot.tex
Fixed reference errors

@@
 %A VPN session between the two hosts is possible, but might be considered as overkill and still does not garantee delivery.
-We have decided to set up a reverse SSH tunnel from the monitoring machine to the honeypot machine, as described in the article ``Remote logging with syslog-ng''. \cite{Remote_logging}
-The command establishes an SSH connection the honeypot, setting options to forego a login shell or anything a normal user would desire, and binds the local socket 9514 on the honeypot machine to tunnel its connections to the same port on our monitoring machine.
-We have put this command there in a script to allow the connection to be re-esthablished quickly and easily if the SSH session is terminated for some reason.
+We have decided to set up a reverse SSH tunnel from the monitoring machine to the honeypot machine, as described in the article ``Remote logging with SSH and syslog-ng''. \cite{remote_logging}
+The example command establishes an SSH connection the honeypot, setting options to forego a login shell or anything a normal user would desire, and binds the local socket 9514 on the honeypot machine to tunnel its connections to the same port on our monitoring machine.
+We have put this command in a script to allow the connection to be re-esthablished quickly and easily if the SSH session is terminated for some reason.
 
-We have configured a locally installed the \verb|syslog-ng| package for our \verb|util| account user, configured it to listen to local TCP connections on port 9514 and thereby receive syslog messages transmitted through this port.
+We have configured a locally installed the \verb|syslog-ng| package for our user on the monitoring machine, and configured it to listen to local TCP connections on port 9514 and thereby receive syslog messages transmitted through this port.
 
@@
 \subsection{scanlogd}
-To detect port scans we have installed the service scanlogd.
+To detect port scans we have installed the service \verb|scanlogd|.
 Per detected scan an entry is written to syslog.
@@
 
-\emph{iptables -I INPUT -m state --state NEW -j LOG --log-prefix "New Connection: "}
-\emph{iptables -I OUTPUT -m state --state NEW -j LOG --log-prefix "New Connection: "}
+\verb|iptables -I INPUT -m state --state NEW -j LOG --log-prefix "New Connection: "|
+\verb|iptables -I OUTPUT -m state --state NEW -j LOG --log-prefix "New Connection: "|
 

### 2011-11-01 Setup_honeypot.tex
added kernel module part

@@
 
+\subsection{Kernel Module}
+
+For HI phase, we would like to have a way to audit the commandline. 
+To make it harder to detect this we would like to do this in the kernel.
+We could edit auditd for the user, but this wil be too noticiable. 
+A tool we can use is Sebek, but this is outdated and not maintained.
+So we have started work on our own kernel module, which will hopefully be ready before we start the HI phase.
+
 \bibliographystyle{alpha}

### 2011-11-01 Setup_honeypot.tex
Auditing/monitoring bit

--- 2011-11-01 Setup_honeypot.tex
@@
 
+\subsection{Monitoring}
+After collecting our syslog information, we want to somehow be alerted of interesting information automatically.
+Some seemingly appropriate projects are available under an open source license, and we have not yet decided which one we want to use.
+A customizable solution is absolutely necessary owing to the wide range of incoming information and the modifications made to a few of the installed software packages specifically for logging to syslog.
+Next to any off-the-shelf solution, some more custom scripting seems inevitable here.
+
  %which makes it possible to use SSH to tunnel the TCP traffic.
@@
 
-To be able to still log on to the honeypot, we configured SSH to use port 9022.
+To be able to still log on to the honeypot, we configured our real SSH service to use port 9022.
+
+We have patched Kippo to log messages to syslog.
 
@@
 \subsection{Snort}
-Snort is an intrusion detection system (IDS) that uses realtime traffic analysis and packet logging on internet protocol networks.
+Snort is an open source intrusion detection system (IDS) that uses realtime traffic analysis and packet logging on internet protocol networks.
 Snort was designed to be used as a network IDS, however we will be running it on our host machine, effectively making it a host-based IDS.
@@
 Unfortunately the VM network went down before we were able to complete our Snort configuration.
-We intend to set up Snort using a predefined rule-set that will only alert on suspicious network behaviour, but do nothing to prevent it. 
-The rules that we will be using are based on a honeypot rule-set. % TODO reference
-Included are rules on detecting backdoors, viruses, DDOS, spam and other abuse.
-These rules will be tweaked if the results are unsatisfactory.
-If a rule is triggered, snort will log an alert.
+We intend to set up Snort using the community rule-set that will only alert on suspicious network behaviour, but do nothing to prevent it. 
+Included are rules on detecting exploit attempts on various common services like HTTP, FTP, SMTP, and many more.
 
-We will configure Snort to send alerts to syslog, and to log its packet captures to some form of read-only filesystem mounted from our monitoring machine (yet to be decided).
-We will install a script like \verb|auto-oink| to automatically retrieve the latest community Snort rules.
+Considering we are in a VM we might be limited in RAM, which Snort seems to require a lot of.
+The configured rules will have to be minimal to avoid exhausting resources.
+
+We will configure Snort to send alerts to syslog, and to log its packet captures to a read-only \verb|sshfs| mounted from our monitoring machine.
+We will install a script like \verb|oinkmaster| to automatically retrieve the latest community Snort rules, with a Snort API key registered specifically for this project.
+
+\subsection{Auditing}
+For the HI phase, we would like to have a way to audit commandline interaction. 
+Various options exist, such as the user-space \verb|auditd| daemon, but to make auditing harder to detect this we feel it would be best to look for a kernel-space solution.
+An auditing module developed by the Honeynet project called Sebek can provide this, but is unmaintained and outdated.
+The grsec kernel patches also provide auditing from kernel space, but we were advised against replacing the stock kernel.
+
+We are still investigating the best way to achieve useful auditing.
+A self-written kernel module that hooks system calls, albeit limited, might suffice to get some much needed auditing information from the honeypot.
 

### 2011-11-01 Setup_honeypot.tex
Typo
More sections

@@
 
-After these period we will analyze the results of both setups and compare the differences in the results we have captured. 
+After these periods we will analyze the results of both setups and compare the differences in the results we have captured. 
 The advantage of a LI honeypot is that the machine can only be compromised to a certain extent and can not really be used for any malicious activity.
@@
 
-\subsection{Syslog}
+\subsection{Syslog-ng}
 By default, \verb|syslog| uses UDP to transmit its log messages to a remote server.
-Relying on UDP is not very desirable, because everything is sent in plaintext and delivery of UDP messages is not guaranteed.
-To allow for safe and reliable transmission of the syslog messages from the honeypot to the monitoring machine, one needs to use TCP connections in syslog, and provide some way of encapsulating the socket used in a cryptographic layer.
+Relying on UDP is not desirable, because everything is sent in plaintext and delivery of UDP messages is not guaranteed.
+To allow for confidential and reliable transmission of the syslog messages from the one machine to another, one needs to use TCP connections in syslog, and provide some way of encapsulating the socket used in a cryptographic layer.
 
@@
 The example command establishes an SSH connection the honeypot, setting options to forego a login shell or anything a normal user would desire, and binds the local socket 9514 on the honeypot machine to tunnel its connections to the same port on our monitoring machine.
-We have put this command in a script to allow the connection to be re-esthablished quickly and easily if the SSH session is terminated for some reason.
 
-We have configured a locally installed the \verb|syslog-ng| package for our user on the monitoring machine, and configured it to listen to local TCP connections on port 9514 and thereby receive syslog messages transmitted through this port.
+We have configured a locally installed the \verb|syslog-ng| package for our user on the monitoring machine, have configured it to listen to local TCP connections on port 9514 and will thereby receive syslog messages transmitted through this port.
+We are aware there is still the possibility of other users on the monitoring machine poisoning our logs, but we feel we can rule out foul play by our peers in this assignment.
 
 To allow for secure SSH access, a public keypair was generated on the util machine and the associated public key was added to the honeypot's \verb|authorized_keys| file.
+
+We have put the command to create the SSH tunnel in a script to allow the connection to be re-esthablished quickly and easily if the SSH session is terminated for some reason. 
+A severance of the connection would lead us to assume the worst initially.
+If re-establishing this connection fails after any attackers have gained access (which should never happen if we monitor adequately), we will have to assume the machine has been compromised beyond our control and will request a shutdown of the VM, if no network downtime was otherwise announced.
 
@@
 
-In this section we decribe the setup of the honeypot.
+In this section we describe the setup of the honeypot.
 
 \subsection{Rsyslog}
-On the honeypot, we've modified the \verb|rsyslog| configuration to also send syslog messages to the (now bound) local port 9514, completing the chain.
-The modified configuration is now stored in its own file in the syslog.d directory, which is a bit obvious to anyone taking interest.
-The configuration should and will be moved over to one of the default configuration files to avoid raising suspicion.
-Note that the chosen port is very similar to the default syslog port 514, and this might give us away.
+On the honeypot, we have modified the default \verb|rsyslog| configuration to also send syslog messages to the (now bound) local port 9514, completing the chain to the syslog instance on our monitoring machine.
+The modified configuration is now stored in its own file in the syslog.d directory, which is a bit obvious to anyone taking an interest.
+We plan to move this small configuration over to one of the default configuration files to avoid raising suspicion outright.
+Note that the chosen port for remote syslog is very similar to the default syslog port 514, and this might give us away.
 We plan to change it to a more nondescript port number.
+
+\subsection{SSHFS mount}
+Not all services support logging to syslog, and for some it is quite infeasible.
+We will create a read-only \verb|sshfs| mount from our monitoring machine that we will use to write our logfiles for various services to.
+We will use interactive login to mount this, as allowing the honeypot back into our monitoring machine with an SSH keypair just seems dangerous.
 
@@
 Kippo is used to log bruteforce attacks and the shell interaction performed by the attacker. 
-To make this as realistic as possible we have added a redirect in the iptables configuration from port 22 to port 2222 that kippo uses. 
+We've installed the latest development version to run as a separate user.
+To make this as realistic as possible we have added a redirect in the iptables configuration from port 22 to port 2222 that kippo uses by default.
 Unfortunately this broke down around the time the firewall was being reconfigured, for unclear reasons. 
@@
 
+This does not log to syslog. We hope to fix that, but it's not a priority until the HI phase.
+
+\section{Apache \& Nginx}
+To attract some web traffic, we have installed the Apache web server, the Apache PHP runtime and the \verb|mod-security| module which allows us to spoof its version information, to make it appear old and vulnerable.
+The Nginx webserver is installed as a reverse proxy in front of Apache, which runs on local port 8080.
+This provides 2 layers of logging (to files), and Nginx's superior transparent caching mechanism allows us to avoid exhausting our resources with scripting languages, should we ever face an application-layer DoS attack (Nginx appears quite resilient against those).
+
+Nginx has been compiled from source and was patched to send the spoofed Apache version header as well.
+We will configure both services to log to our \verb|sshfs| mount.
+
 \subsection{Google Hack Honeypot}
@@
 
-To set up GHH we have first installed Apache and PHP5.
+%To set up GHH we have first installed Apache and PHP5.
 We configured GHH to log to in CSV format, since it requires minimal setup and is easily parsed.
@@
 
-We will configure Snort to send alerts to syslog, and to log its packet captures to a read-only \verb|sshfs| mounted from our monitoring machine.
+We will configure Snort to send alerts to syslog, and to log its packet captures to the read-only \verb|sshfs| mounted from our monitoring machine.
 We will install a script like \verb|oinkmaster| to automatically retrieve the latest community Snort rules, with a Snort API key registered specifically for this project.

### 2011-11-01 Setup_honeypot.tex
Small additions

@@
 
-\verb|iptables -I INPUT -m state --state NEW -j LOG --log-prefix "New Connection: "|
-\verb|iptables -I OUTPUT -m state --state NEW -j LOG --log-prefix "New Connection: "|
+\begin{verbatim}
+iptables -I INPUT -m state --state NEW -j LOG --log-prefix "New Connection: "
+iptables -I OUTPUT -m state --state NEW -j LOG --log-prefix "New Connection: "
+\end{verbatim}
 
@@
 To capture attackers that try to come in by brute-forcing SSH logins, we are using Kippo. 
-Kippo is used to log bruteforce attacks and the shell interaction performed by the attacker. 
+Kippo is a sandbox SSH service, used to log bruteforce attacks and the shell interaction performed by the attacker after gaining access (with a trivial password). 
 We've installed the latest development version to run as a separate user.
-To make this as realistic as possible we have added a redirect in the iptables configuration from port 22 to port 2222 that kippo uses by default.
-Unfortunately this broke down around the time the firewall was being reconfigured, for unclear reasons. 
-The reason for forwarding was not to give Kippo root capabilities.
 
-To be able to still log on to the honeypot, we configured our real SSH service to use port 9022.
+To make this as realistic as possible we have added a redirect in the iptables configuration from the real SSH port 22 to port 2222 that Kippo uses by default.
+The reason for forwarding in iptables was to not give Kippo root rights, which would be very dangerous. % better word: insane
+Unfortunately the forward broke down around the time the firewall was being reconfigured, for reasons yet unclear . 
+If it can't be resolved, we can circumvent this issue by running Kippo through a copy of the Python interpreter that has the capability to bind privileged ports.
+
+To be able to still log on to the honeypot, we have configured our real SSH service to run on port 9022.
 
@@
 \subsection{DenyHosts}
-If someone finds out our real SSH port changed to 9022, they might try to break in via our real SSH service.
+If someone finds out our real SSH port changed to 9022, they might try to break in through our real SSH service.
 To make sure no bruteforce attacks can be done on this port, DenyHosts is installed.
-This service blocks IP-adresses that try to log in more than 3 times with the wrong password by adding them to the \verb|hosts.deny| file.
+This service blocks IP-addresses that try to log in more than 3 times with the wrong password by adding them to the \verb|hosts.deny| file.
 The hosts listed there are blacklisted on the machine, but we can edit them if needed.

### 2011-11-01 Setup_honeypot.tex
Fixed denyhosts onzin
added wordpress and editted iptables

@@
 
+We are also going to filter outgoing SMTP, SIP and NetBIOS.
+
 \subsection{Kippo}
@@
 The hosts listed there are blacklisted on the machine, but we can edit them if needed.
-
-This does not log to syslog. We hope to fix that, but it's not a priority until the HI phase.
+Denyhosts can log to syslog by editing its configuration file.
 
@@
 Then, we patched the GHH code to allow logging to syslog, to maintain a single, consistent point of log collection.
+
+We also hope to detect and capture a bruteforce attack or scan on generic filenames, for example \verb|passwd.txt|. 
+This will probably warn us that an attacker will try to use a dictionary attack.
+
+\subsection{Wordpress}
+
+To capture more attackers during the HI phase, we plan to use an outdated version of WordPress. 
+Wordpress is software, powered by PHP and MySQL, that can be used to easily create websites or blogs and manage them via the content management system. 
+Unfortunately exploits are frequently found for Wordpress and patches are published on a regular basis. 
+We plan to use an older version of Wordpress, for example a version below version 3.0.4, to be sure that we use a version with some exploits.
 

### 2011-11-01 Setup_honeypot.tex
added some small changes

@@
 
-In this document we explain how we have set up the our honeypot. 
+In this document we explain how we have set up the honeypot. 
 We have decided to split up our monitoring time into two parts.
@@
 
-We are also going to filter outgoing SMTP, SIP and NetBIOS.
+We are also going to filter outgoing SMTP, SIP and NetBIOS traffic.
 
@@
 Wordpress is software, powered by PHP and MySQL, that can be used to easily create websites or blogs and manage them via the content management system. 
-Unfortunately exploits are frequently found for Wordpress and patches are published on a regular basis. 
-We plan to use an older version of Wordpress, for example a version below version 3.0.4, to be sure that we use a version with some exploits.
+Exploits are found on a frequent basis for Wordpress and patches are published on a regular basis. 
+We plan to use an older version of Wordpress, for example a version below version 3.0.4, to be sure that we use a version for which public exploits exist.
 

### 2012-01-05 Paper_honeypot.tex
Opzet honeypot paper

\documentclass[11pt]{article}
\usepackage{a4wide, graphicx, fancyhdr, wrapfig, tabularx, amsmath, amssymb, hyperref, color, verbatim, nameref}
\usepackage[english]{babel}
\definecolor{linkcolour}{rgb}{0,0.2,0.6}
\hypersetup{colorlinks,breaklinks,urlcolor=linkcolour, linkcolor=linkcolour}

%----------------------- Macros and Definitions --------------------------

\setlength\headheight{20pt}\usepackage{}
\addtolength\topmargin{-10pt}
%\addtolength\footskip{20pt}

\fancypagestyle{plain}{%
\fancyhf{}
\fancyfoot[RO,LE]{\sffamily\bfseries\thepage}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}
}

\pagestyle{fancy}
\fancyhf{}
\fancyfoot[RO,LE]{\sffamily\bfseries\thepage}
\fancyhead[RO,LE]{\textsc{}}
\fancyhead[LO,RE]{\emph{}}
\renewcommand{\headrulewidth}{1pt}
\renewcommand{\footrulewidth}{0pt}
\newcommand{\tab}{\hspace*{2em}}

\newcommand{\tocheck}[1]{{\bf !?: #1 :!?}}
\newcommand{\OBA}{Online behavioural advertising }
\newcommand{\oba}{online behavioural advertising }
\newcommand{\ePD}{ePrivacy Directive }
\newcommand{\DPD}{Data Protection Directive }

\frenchspacing

%-------------------------------- Title ----------------------------------

\title{\textbf{Law in Cyberspace\\ \emph{Web Analytics and Online Personalisation}}}
\author{Aram Verstegen, \tocheck{studentnummer Utwente} \\
	 Mark Vijfvinkel, 1275194 \\
	 Bart Lutgens, \tocheck{studentnummer Utwente} \\
	Joost Kremers, \tocheck{studentnummer Utwente}}

\date{\today}

\begin{document}
\maketitle

\section{Introduction}

\section{Setup}
\label{Setup}

On the 1st of November we delivered our initial setup document, however during the "Waiting for hackers" phase we have made adjustments to the setup in order to attract more activity to the honeypot. In this section we will discuss the changes that were made.

\section{Statistics}
\label{Statistics}
In this section we will discuss the statistical data that have been gathered from our logfiles.

\section{Attacks on the different services}
\label{Attacks}

\section{Connection between the attacks}
\label{Connection}

\section{Conclusion}
\label{Conclusion}

%%% REFS %%%

\bibliographystyle{plain} % amsalpha
\bibliography{assignment2}

\end{document}

### 2012-01-05 Paper_honeypot.tex
Intro toegevoegd

@@
 \newcommand{\tocheck}[1]{{\bf !?: #1 :!?}}
-\newcommand{\OBA}{Online behavioural advertising }
-\newcommand{\oba}{online behavioural advertising }
-\newcommand{\ePD}{ePrivacy Directive }
-\newcommand{\DPD}{Data Protection Directive }
+%\newcommand{\OBA}{Online behavioural advertising }
 
@@
 
-\title{\textbf{Law in Cyberspace\\ \emph{Web Analytics and Online Personalisation}}}
+\title{\textbf{Honeypot Assignment: \\ \emph{Group Contagiarium}}}
 \author{Aram Verstegen, \tocheck{studentnummer Utwente} \\
@@
 \section{Introduction}
+In this paper we will try to present an overview of the data we have collected from the honeypot and explain the different attacks that were waged against it. 
+In the second section we will describe the changes we have made to the setup during the "Waiting for hackers" phase. 
+The third section will present an overview of various statistics we have collected from the logfiles, e.g. number of IP-addresses, countries and attack types. 
+In the fourth section we will discuss which attacks took place against the different services the honeypot was running e.g. SSH and MySQL.
+In order to discover the sophisticationof the attacks, we will try to present possible connections between an attack on one service and an attack on a different service, in the fifth section.
+The sixth section will conclude our findings. 
+Finally we will discuss this project and provide some improvements.
 
@@
 \label{Attacks}
+During phase 2 we have monitored the honeypot closely and collected data about different attacks against the honeypot.
+In this section we will discuss the attacks against the services the honeypot was running.
 
@@
 \label{Connection}
+In this section we will correlate the data about attacks against one service against the data of attacks of a different service.
+We will try to look for a pattern or similarities in this data and determine if the attack was more sophisticated.
 
@@
 \label{Conclusion}
+From Section \ref{Connection} we can conclude that...
+
+\section{Discussion}
+\label{Discussion}
+In this section we would like to discuss some issues we encountered during the honeypotproject.
+
+\begin{itemize}
+\item Low/High interaction phase
+\item GHH
+\item etc
+\end{itemize}
 

### 2012-01-05 Paper_honeypot.tex
setup toegevoegd

@@
 \section{Introduction}
+\tocheck{Kijken of the secties nog kloppen!}
 In this paper we will try to present an overview of the data we have collected from the honeypot and explain the different attacks that were waged against it. 
@@
 On the 1st of November we delivered our initial setup document, however during the "Waiting for hackers" phase we have made adjustments to the setup in order to attract more activity to the honeypot.
-In this section we will discuss the changes that were made.
+In this section we will present the initial setup and in addittion to that, discuss the changes that were made.
+
+%%%%MOET NOG WORDEN HERZIEN%%%%%
+
+\section{Setup: Monitoring Machine}
+
+In this section we describe the setup of the monitoring machine.
+
+\subsection{Syslog-ng}
+By default, \verb|syslog| uses UDP to transmit its log messages to a remote server.
+Relying on UDP is not desirable, because everything is sent in plaintext and delivery of UDP messages is not guaranteed.
+To allow for confidential and reliable transmission of the syslog messages from the one machine to another, one needs to use TCP connections in syslog, and provide some way of encapsulating the socket used in a cryptographic layer.
+
+We have decided to set up a reverse SSH tunnel from the monitoring machine to the honeypot machine, as described in the article ``Remote logging with SSH and syslog-ng''. \cite{remote_logging}
+The example command establishes an SSH connection the honeypot, setting options to forego a login shell or anything a normal user would desire, and binds the local socket 9514 on the honeypot machine to tunnel its connections to the same port on our monitoring machine.
+
+We have configured a locally installed the \verb|syslog-ng| package for our user on the monitoring machine, have configured it to listen to local TCP connections on port 9514 and will thereby receive syslog messages transmitted through this port.
+We are aware there is still the possibility of other users on the monitoring machine poisoning our logs, but we feel we can rule out foul play by our peers in this assignment.
+
+To allow for secure SSH access, a public keypair was generated on the util machine and the associated public key was added to the honeypot's \verb|authorized_keys| file.
+
+We have put the command to create the SSH tunnel in a script to allow the connection to be re-esthablished quickly and easily if the SSH session is terminated for some reason. 
+A severance of the connection would lead us to assume the worst initially.
+If re-establishing this connection fails after any attackers have gained access (which should never happen if we monitor adequately), we will have to assume the machine has been compromised beyond our control and will request a shutdown of the VM, if no network downtime was otherwise announced.
+
+\subsection{Monitoring}
+After collecting our syslog information, we want to somehow be alerted of interesting information automatically.
+Some seemingly appropriate projects are available under an open source license, and we have not yet decided which one we want to use.
+A customizable solution is absolutely necessary owing to the wide range of incoming information and the modifications made to a few of the installed software packages specifically for logging to syslog.
+Next to any off-the-shelf solution, some more custom scripting seems inevitable here.
+
+\section{Setup: Honeypot}
+
+In this section we describe the setup of the honeypot.
+
+\subsection{Rsyslog}
+On the honeypot, we have modified the default \verb|rsyslog| configuration to also send syslog messages to the (now bound) local port 9514, completing the chain to the syslog instance on our monitoring machine.
+The modified configuration is now stored in its own file in the syslog.d directory, which is a bit obvious to anyone taking an interest.
+We plan to move this small configuration over to one of the default configuration files to avoid raising suspicion outright.
+Note that the chosen port for remote syslog is very similar to the default syslog port 514, and this might give us away.
+We plan to change it to a more nondescript port number.
+
+\subsection{SSHFS mount}
+Not all services support logging to syslog, and for some it is quite infeasible.
+We will create a read-only \verb|sshfs| mount from our monitoring machine that we will use to write our logfiles for various services to.
+We will use interactive login to mount this, as allowing the honeypot back into our monitoring machine with an SSH keypair just seems dangerous.
+
+\subsection{scanlogd}
+To detect port scans we have installed the service \verb|scanlogd|.
+Per detected scan an entry is written to syslog.
+
+\subsection{iptables}
+To log all incoming and outgoing connection attempts, we have added the following rules to the iptables configuration. 
+This will also log to syslog.
+
+\begin{verbatim}
+iptables -I INPUT -m state --state NEW -j LOG --log-prefix "New Connection: "
+iptables -I OUTPUT -m state --state NEW -j LOG --log-prefix "New Connection: "
+\end{verbatim}
+
+We are also going to filter outgoing SMTP, SIP and NetBIOS traffic.
+
+\subsection{Kippo}
+To capture attackers that try to come in by brute-forcing SSH logins, we are using Kippo. 
+Kippo is a sandbox SSH service, used to log bruteforce attacks and the shell interaction performed by the attacker after gaining access (with a trivial password). 
+We've installed the latest development version to run as a separate user.
+
+To make this as realistic as possible we have added a redirect in the iptables configuration from the real SSH port 22 to port 2222 that Kippo uses by default.
+The reason for forwarding in iptables was to not give Kippo root rights, which would be very dangerous.
+Unfortunately the forward broke down around the time the firewall was being reconfigured, for reasons yet unclear . 
+If it can't be resolved, we can circumvent this issue by running Kippo through a copy of the Python interpreter that has the capability to bind privileged ports.
+
+To be able to still log on to the honeypot, we have configured our real SSH service to run on port 9022.
+
+We have patched Kippo to log messages to syslog.
+
+\subsection{DenyHosts}
+If someone finds out our real SSH port changed to 9022, they might try to break in through our real SSH service.
+To make sure no bruteforce attacks can be done on this port, DenyHosts is installed.
+This service blocks IP-addresses that try to log in more than 3 times with the wrong password by adding them to the \verb|hosts.deny| file.
+The hosts listed there are blacklisted on the machine, but we can edit them if needed.
+Denyhosts can log to syslog by editing its configuration file.
+
+\section{Apache \& Nginx}
+To attract some web traffic, we have installed the Apache web server, the Apache PHP runtime and the \verb|mod-security| module which allows us to spoof its version information, to make it appear old and vulnerable.
+The Nginx webserver is installed as a reverse proxy in front of Apache, which runs on local port 8080.
+This provides 2 layers of logging (to files), and Nginx's superior transparent caching mechanism allows us to avoid exhausting our resources with scripting languages, should we ever face an application-layer DoS attack (Nginx appears quite resilient against those).
+
+Nginx has been compiled from source and was patched to send the spoofed Apache version header as well.
+We will configure both services to log to our \verb|sshfs| mount.
+
+\subsection{Google Hack Honeypot}
+Google Hack Honeypot (GHH) is designed to be used to attract search engine attackers. 
+These attackers will use search queries to find sensitive data, for example password files, that due to some misconfiguration ended up being indexed by a search engine like Google.
+GHH allows logging of these requests to a MySQL database or Comma Seperated Value (CSV) file.
+
+Since GHH offers multiple search targets, we picked one to start with, namely GGH version 1.2 passlist.txt.
+We had no particular reason for this, only that it was easy to set up. 
+We plan to install more variants of the GHH to attract more attackers, but before this could happen the VM's were shut down.
+
+%To set up GHH we have first installed Apache and PHP5.
+We configured GHH to log to in CSV format, since it requires minimal setup and is easily parsed.
+%We can later reconfigure to use MySQL, if the results are not satisfactory.
+Then, we patched the GHH code to allow logging to syslog, to maintain a single, consistent point of log collection.
+
+We also hope to detect and capture a bruteforce attack or scan on generic filenames, for example \verb|passwd.txt|. 
+This will probably warn us that an attacker will try to use a dictionary attack.
+
+\subsection{Wordpress}
+
+To capture more attackers during the HI phase, we plan to use an outdated version of WordPress. 
+Wordpress is software, powered by PHP and MySQL, that can be used to easily create websites or blogs and manage them via the content management system. 
+Exploits are found on a frequent basis for Wordpress and patches are published on a regular basis. 
+We plan to use an older version of Wordpress, for example a version below version 3.0.4, to be sure that we use a version for which public exploits exist.
+
+\subsection{Snort}
+Snort is an open source intrusion detection system (IDS) that uses realtime traffic analysis and packet logging on internet protocol networks.
+Snort was designed to be used as a network IDS, however we will be running it on our host machine, effectively making it a host-based IDS.
+
+Unfortunately the VM network went down before we were able to complete our Snort configuration.
+We intend to set up Snort using the community rule-set that will only alert on suspicious network behaviour, but do nothing to prevent it. 
+Included are rules on detecting exploit attempts on various common services like HTTP, FTP, SMTP, and many more.
+
+Considering we are in a VM we might be limited in RAM, which Snort seems to require a lot of.
+The configured rules will have to be minimal to avoid exhausting resources.
+
+We will configure Snort to send alerts to syslog, and to log its packet captures to the read-only \verb|sshfs| mounted from our monitoring machine.
+We will install a script like \verb|oinkmaster| to automatically retrieve the latest community Snort rules, with a Snort API key registered specifically for this project.
+
+\subsection{Auditing}
+For the HI phase, we would like to have a way to audit commandline interaction. 
+Various options exist, such as the user-space \verb|auditd| daemon, but to make auditing harder to detect this we feel it would be best to look for a kernel-space solution.
+An auditing module developed by the Honeynet project called Sebek can provide this, but is unmaintained and outdated.
+The grsec kernel patches also provide auditing from kernel space, but we were advised against replacing the stock kernel.
+
+We are still investigating the best way to achieve useful auditing.
+A self-written kernel module that hooks system calls, albeit limited, might suffice to get some much needed auditing information from the honeypot.
 

### 2012-01-09 Paper_honeypot.tex
setup gedeeltelijk bijgewerkt

@@
 
-\section{Setup: Monitoring Machine}
+\subsection{Monitoring Machine}
 
-In this section we describe the setup of the monitoring machine.
+In this subsection we describe the setup of the monitoring machine.
 
-\subsection{Syslog-ng}
+\subsubsection{Syslog-ng}
 By default, \verb|syslog| uses UDP to transmit its log messages to a remote server.
@@
 
-We have decided to set up a reverse SSH tunnel from the monitoring machine to the honeypot machine, as described in the article ``Remote logging with SSH and syslog-ng''. \cite{remote_logging}
+We have decided to set up a reverse SSH tunnel from the monitoring machine to the honeypot machine, as described in the article ``Remote logging with SSH and syslog-ng''.
+\cite{remote_logging}\tocheck{Toevoegen refs}
 The example command establishes an SSH connection the honeypot, setting options to forego a login shell or anything a normal user would desire, and binds the local socket 9514 on the honeypot machine to tunnel its connections to the same port on our monitoring machine.
 
-We have configured a locally installed the \verb|syslog-ng| package for our user on the monitoring machine, have configured it to listen to local TCP connections on port 9514 and will thereby receive syslog messages transmitted through this port.
-We are aware there is still the possibility of other users on the monitoring machine poisoning our logs, but we feel we can rule out foul play by our peers in this assignment.
+We have configured a locally installed \verb|syslog-ng| package for our user on the monitoring machine and have configured it to listen to local TCP connections on port 9514 and will thereby receive syslog messages transmitted through this port.
+%We are aware there is still the possibility of other users on the monitoring machine poisoning our logs, but we feel we can rule out foul play by our peers in this assignment.
 
@@
 A severance of the connection would lead us to assume the worst initially.
-If re-establishing this connection fails after any attackers have gained access (which should never happen if we monitor adequately), we will have to assume the machine has been compromised beyond our control and will request a shutdown of the VM, if no network downtime was otherwise announced.
+If re-establishing this connection fails after any attackers have gained access (which should never happen if we monitor adequately), we would have assumed that the machine was compromised beyond our control and would have requested a shutdown of the VM, if no network downtime was otherwise announced.
 
-\subsection{Monitoring}
-After collecting our syslog information, we want to somehow be alerted of interesting information automatically.
-Some seemingly appropriate projects are available under an open source license, and we have not yet decided which one we want to use.
-A customizable solution is absolutely necessary owing to the wide range of incoming information and the modifications made to a few of the installed software packages specifically for logging to syslog.
-Next to any off-the-shelf solution, some more custom scripting seems inevitable here.
+\subsubsection{Monitoring}
+After collecting our syslog information, we wanted to be somehow alerted of interesting information automatically.
+We decided to use an IRC bot that would leave a message in an IRC channel (\#contagiarium) everytime an entry was made to the syslog file.
+First this lead to an overloaded channel and several false positives.
+To solve these problems filters were created to only post relevant information.
+For example, from the kippo logs only the actual login attempts would be shown with the tried password.
 
-\section{Setup: Honeypot}
+\subsection{Honeypot}
 
@@
 
-\subsection{Rsyslog}
-On the honeypot, we have modified the default \verb|rsyslog| configuration to also send syslog messages to the (now bound) local port 9514, completing the chain to the syslog instance on our monitoring machine.
+\subsubsection{Rsyslog}
+On the honeypot, we modified the default \verb|rsyslog| configuration to also send syslog messages to the (now bound) local port 9514, completing the chain to the syslog instance on our monitoring machine.
 The modified configuration is now stored in its own file in the syslog.d directory, which is a bit obvious to anyone taking an interest.
+%%Bespreken Aram
 We plan to move this small configuration over to one of the default configuration files to avoid raising suspicion outright.
@@
 
-\subsection{SSHFS mount}
-Not all services support logging to syslog, and for some it is quite infeasible.
-We will create a read-only \verb|sshfs| mount from our monitoring machine that we will use to write our logfiles for various services to.
-We will use interactive login to mount this, as allowing the honeypot back into our monitoring machine with an SSH keypair just seems dangerous.
+\subsubsection{SSHFS mount}
+Not all services support logging to syslog, and for some it is infeasible.
+We created a read-only \verb|sshfs| mount from our monitoring machine that we will use to write our logfiles for various services to.
+We used interactive login to mount this, as allowing the honeypot back into our monitoring machine with an SSH keypair just seems dangerous.
 
-\subsection{scanlogd}
+\subsubsection{scanlogd}
 To detect port scans we have installed the service \verb|scanlogd|.
@@
 
-\subsection{iptables}
-To log all incoming and outgoing connection attempts, we have added the following rules to the iptables configuration. 
-This will also log to syslog.
+\subsubsection{iptables}
+To log all incoming and outgoing connection attempts, we have added the following rules to the iptables configuration. The attempts were also logged to syslog.
 
@@
 
-We are also going to filter outgoing SMTP, SIP and NetBIOS traffic.
+%Bespreken Aram
+%We are also going to filter outgoing SMTP, SIP and NetBIOS traffic.
 
-\subsection{Kippo}
-To capture attackers that try to come in by brute-forcing SSH logins, we are using Kippo. 
+\subsubsection{Kippo}
+To capture attackers that try to come in by brute-forcing SSH logins, we are used Kippo. 
 Kippo is a sandbox SSH service, used to log bruteforce attacks and the shell interaction performed by the attacker after gaining access (with a trivial password). 
-We've installed the latest development version to run as a separate user.
+The latest development version was installed to run as a separate user.
 
-To make this as realistic as possible we have added a redirect in the iptables configuration from the real SSH port 22 to port 2222 that Kippo uses by default.
+To make this as realistic as possible we added a redirect in the iptables configuration from the real SSH port 22 to port 2222 that Kippo uses by default.
 The reason for forwarding in iptables was to not give Kippo root rights, which would be very dangerous.
+%Bespreken Aram
 Unfortunately the forward broke down around the time the firewall was being reconfigured, for reasons yet unclear . 
@@
 
-To be able to still log on to the honeypot, we have configured our real SSH service to run on port 9022.
+To be able to still log on to the honeypot, we configured our real SSH service to run on port 9022.
 
-We have patched Kippo to log messages to syslog.
+Furthermore, we patched Kippo to log messages to syslog.
 
-\subsection{DenyHosts}
-If someone finds out our real SSH port changed to 9022, they might try to break in through our real SSH service.
-To make sure no bruteforce attacks can be done on this port, DenyHosts is installed.
+During the second phase of the project, we noticed that of the succesfull bruteforce attempts against Kippo, none actually tried to continue and explore or use the honeypot.
+Running a default portscan with Metasploit revealed that it can detect if Kippo is being used.
+We thought this explained why nobody tried to log in.
+To solve this problem, we adjusted the Kippo configuration so that Metasploit could recognize Kippo.
+Unfortunately did this not lead to actual exploration or use of the honeypot.
+
+\subsubsection{DenyHosts}
+If someone found out that our real SSH port was changed to 9022, they might try to break in through the real SSH service.
+To make sure no bruteforce attacks can be done on this port, DenyHosts was installed.
 This service blocks IP-addresses that try to log in more than 3 times with the wrong password by adding them to the \verb|hosts.deny| file.
 The hosts listed there are blacklisted on the machine, but we can edit them if needed.
+%bespreken Aram
 Denyhosts can log to syslog by editing its configuration file.
 
-\section{Apache \& Nginx}
-To attract some web traffic, we have installed the Apache web server, the Apache PHP runtime and the \verb|mod-security| module which allows us to spoof its version information, to make it appear old and vulnerable.
+\subsection{Apache \& Nginx}
+To attract web traffic, we installed the Apache web server, the Apache PHP runtime and the \verb|mod-security| module which allows us to spoof its version information, to make it appear old and vulnerable.
 The Nginx webserver is installed as a reverse proxy in front of Apache, which runs on local port 8080.
-This provides 2 layers of logging (to files), and Nginx's superior transparent caching mechanism allows us to avoid exhausting our resources with scripting languages, should we ever face an application-layer DoS attack (Nginx appears quite resilient against those).
+This provides 2 layers of logging (to files), and Nginx's superior transparent caching mechanism allows us to avoid exhausting our resources with scripting languages, should we have ever faced an application-layer DoS attack (Nginx appears quite resilient against those).
 
@@
 
-\subsection{Google Hack Honeypot}
+\subsubsection{Google Hack Honeypot}
 Google Hack Honeypot (GHH) is designed to be used to attract search engine attackers. 
@@
 We had no particular reason for this, only that it was easy to set up. 
-We plan to install more variants of the GHH to attract more attackers, but before this could happen the VM's were shut down.
+%We plan to install more variants of the GHH to attract more attackers, but before this could happen the VM's were shut down.
 
@@
 
-We also hope to detect and capture a bruteforce attack or scan on generic filenames, for example \verb|passwd.txt|. 
-This will probably warn us that an attacker will try to use a dictionary attack.
+We hoped to detect and capture a bruteforce attack or scan on generic filenames, for example \verb|passwd.txt|, because this would probably have warned us that an attacker will try to use a dictionary attack.
+Unfortunately GGH broke down after the Virtual Machines were shut down, we were unable to get it back running and therefore decided not invest more time and effort into it.
+We also believe it would not have brought us more interesting data.
+
+\subsubsection{Website}
 

### 2012-01-09 Paper_honeypot.tex
setup verder uitgebreid

@@
 We thought this explained why nobody tried to log in.
-To solve this problem, we adjusted the Kippo configuration so that Metasploit could recognize Kippo.
+To solve this problem, we adjusted the Kippo configuration so that Metasploit could not recognize Kippo.
 Unfortunately did this not lead to actual exploration or use of the honeypot.
@@
 \subsubsection{Website}
+To attract more activity we decided to create a fake website that would resemble a website of the University of Twente.
+To make it more appealing in ways of computing power we used the department of Partical Physics and Astronomy as a cover.
+On the website we also put some information about a fictive person (Jenny Abott) would manage the servers and website.
+So we created a useraccount with the same login and password, namely Jenny.
+This way a more sophisticated attack would be possible.
+An attacker would look at the website, determine if it is intresing to attack and then connect to Kippo to log in as Jenny.
 
-\subsection{Wordpress}
+\subsubsection{Wordpress}
+In order to make the website more vunerable and appealing to hackers, we installed an out of date version of Wordpress.
+Wordpress is software, powered by PHP and MySQL, that can be used to easily create websites or blogs and manage them via the content management system.
+Exploits are found on a frequent basis for Wordpress and therefore patches are published frequently.
+In the first week of December we installed version 2.8.4 of Wordpress on the honeypot.
 
-To capture more attackers during the HI phase, we plan to use an outdated version of WordPress. 
-Wordpress is software, powered by PHP and MySQL, that can be used to easily create websites or blogs and manage them via the content management system. 
-Exploits are found on a frequent basis for Wordpress and patches are published on a regular basis. 
-We plan to use an older version of Wordpress, for example a version below version 3.0.4, to be sure that we use a version for which public exploits exist.
-
-\subsection{Snort}
+\subsection{Snort and Suricata}
 Snort is an open source intrusion detection system (IDS) that uses realtime traffic analysis and packet logging on internet protocol networks.
@@
 
-Unfortunately the VM network went down before we were able to complete our Snort configuration.
-We intend to set up Snort using the community rule-set that will only alert on suspicious network behaviour, but do nothing to prevent it. 
+We set up Snor using the community rule-set that will only alert on suspicious network behaviour, but do nothing to prevent it. 
 Included are rules on detecting exploit attempts on various common services like HTTP, FTP, SMTP, and many more.
 
-Considering we are in a VM we might be limited in RAM, which Snort seems to require a lot of.
-The configured rules will have to be minimal to avoid exhausting resources.
+We configured Snort to send alerts to syslog, and to log its packet captures to the read-only \verb|sshfs| mounted from our monitoring machine.
+We also installed \verb|oinkmaster| to automatically retrieve the latest community Snort rules, with a Snort API key registered specifically for this project.
 
-We will configure Snort to send alerts to syslog, and to log its packet captures to the read-only \verb|sshfs| mounted from our monitoring machine.
-We will install a script like \verb|oinkmaster| to automatically retrieve the latest community Snort rules, with a Snort API key registered specifically for this project.
+Unfortunately we noticed that Snort required a lot of resources and was slowing down the honeypot. Instead of Snort we decided to use Suricata, which does the same thing, but does it more efficiently.
+
+%Considering we are in a VM we might be limited in RAM, which Snort seems to require a lot of.
+%The configured rules will have to be minimal to avoid exhausting resources.
 
@@
 \bibliographystyle{plain} % amsalpha
-\bibliography{assignment2}
+\bibliography{ref_honeypot}
 

### 2012-01-09 Paper_honeypot.tex
Setup aangepast!

@@
 	 Mark Vijfvinkel, 1275194 \\
-	 Bart Lutgens, \tocheck{studentnummer Utwente} \\
-	Joost Kremers, \tocheck{studentnummer Utwente}}
+	 Bart Lutgens, 12752408 \\
+	Joost Kremers, 1275259}
 
@@
 
-We have decided to set up a reverse SSH tunnel from the monitoring machine to the honeypot machine, as described in the article ``Remote logging with SSH and syslog-ng''.
-\cite{remote_logging}\tocheck{Toevoegen refs}
+We decided to set up a reverse SSH tunnel from the monitoring machine to the honeypot machine, as described in the article "Remote logging with SSH and syslog-ng''. \cite{remote_logging}
 The example command establishes an SSH connection the honeypot, setting options to forego a login shell or anything a normal user would desire, and binds the local socket 9514 on the honeypot machine to tunnel its connections to the same port on our monitoring machine.
 
-We have configured a locally installed \verb|syslog-ng| package for our user on the monitoring machine and have configured it to listen to local TCP connections on port 9514 and will thereby receive syslog messages transmitted through this port.
+We configured a locally installed \verb|syslog-ng| package for our user on the monitoring machine and have configured it to listen to local TCP connections on port 9514 and will thereby receive syslog messages transmitted through this port.
 %We are aware there is still the possibility of other users on the monitoring machine poisoning our logs, but we feel we can rule out foul play by our peers in this assignment.
@@
 First this lead to an overloaded channel and several false positives.
-To solve these problems filters were created to only post relevant information.
+To solve these problems filters (regular expressions in a file) were created to only post relevant information.
 For example, from the kippo logs only the actual login attempts would be shown with the tried password.
@@
 On the honeypot, we modified the default \verb|rsyslog| configuration to also send syslog messages to the (now bound) local port 9514, completing the chain to the syslog instance on our monitoring machine.
-The modified configuration is now stored in its own file in the syslog.d directory, which is a bit obvious to anyone taking an interest.
-%%Bespreken Aram
-We plan to move this small configuration over to one of the default configuration files to avoid raising suspicion outright.
-Note that the chosen port for remote syslog is very similar to the default syslog port 514, and this might give us away.
-We plan to change it to a more nondescript port number.
+At first the modified configuration was stored in its own file in the syslog.d directory, which is a bit obvious to anyone taking an interest.
+Therefore we moved the configuration to the standard syslog directory.
+
+%We plan to move this small configuration over to one of the default configuration files to avoid raising suspicion outright.
+Note that the chosen port for remote syslog is very similar to the default syslog port 514, which might have given us away. So we changed the port number to 9514.
+%We plan to change it to a more nondescript port number.
 
 \subsubsection{SSHFS mount}
-Not all services support logging to syslog, and for some it is infeasible.
-We created a read-only \verb|sshfs| mount from our monitoring machine that we will use to write our logfiles for various services to.
-We used interactive login to mount this, as allowing the honeypot back into our monitoring machine with an SSH keypair just seems dangerous.
+Not all services support logging to syslog and for some it is infeasible.
+We created a read-only \verb|sshfs| mount from our monitoring machine that we used to write our logfiles for various services to.
+We used an interactive login to mount this, as allowing the honeypot back into our monitoring machine with an SSH keypair just seems dangerous.
 
 \subsubsection{scanlogd}
-To detect port scans we have installed the service \verb|scanlogd|.
+To detect port scans we installed the service \verb|scanlogd|.
 Per detected scan an entry is written to syslog.
@@
 
-%Bespreken Aram
+We have filtered outgoing SMTP, SIP and NetBIOS traffic, in order to prevent abuse and spam.
 %We are also going to filter outgoing SMTP, SIP and NetBIOS traffic.
@@
 \subsubsection{Kippo}
-To capture attackers that try to come in by brute-forcing SSH logins, we are used Kippo. 
+To capture attackers that try to come in by brute-forcing SSH logins, we used Kippo. 
 Kippo is a sandbox SSH service, used to log bruteforce attacks and the shell interaction performed by the attacker after gaining access (with a trivial password). 
@@
 The reason for forwarding in iptables was to not give Kippo root rights, which would be very dangerous.
-%Bespreken Aram
-Unfortunately the forward broke down around the time the firewall was being reconfigured, for reasons yet unclear . 
-If it can't be resolved, we can circumvent this issue by running Kippo through a copy of the Python interpreter that has the capability to bind privileged ports.
+
+%Unfortunately the forward broke down around the time the firewall was being reconfigured, for reasons yet unclear . 
+%If it can't be resolved, we can circumvent this issue by running Kippo through a copy of the Python interpreter that has the capability to bind privileged ports.
 
@@
 To solve this problem, we adjusted the Kippo configuration so that Metasploit could not recognize Kippo.
-Unfortunately did this not lead to actual exploration or use of the honeypot.
+Unfortunately this did still not lead to actual exploration or use of the honeypot.
 
@@
 If someone found out that our real SSH port was changed to 9022, they might try to break in through the real SSH service.
-To make sure no bruteforce attacks can be done on this port, DenyHosts was installed.
+To make sure no bruteforce attacks could be done on this port, DenyHosts was installed.
 This service blocks IP-addresses that try to log in more than 3 times with the wrong password by adding them to the \verb|hosts.deny| file.
 The hosts listed there are blacklisted on the machine, but we can edit them if needed.
-%bespreken Aram
-Denyhosts can log to syslog by editing its configuration file.
+
+%Denyhosts can log to syslog by editing its configuration file.
 
@@
 We hoped to detect and capture a bruteforce attack or scan on generic filenames, for example \verb|passwd.txt|, because this would probably have warned us that an attacker will try to use a dictionary attack.
-Unfortunately GGH broke down after the Virtual Machines were shut down, we were unable to get it back running and therefore decided not invest more time and effort into it.
+Unfortunately GGH broke down after the Virtual Machines were shut down, we were unable to get it back up running and therefore decided not invest more time and effort into it.
 We also believe it would not have brought us more interesting data.
@@
 
-\subsection{Snort and Suricata}
+\subsubsection{Snort and Suricata}
 Snort is an open source intrusion detection system (IDS) that uses realtime traffic analysis and packet logging on internet protocol networks.
@@
 
-Unfortunately we noticed that Snort required a lot of resources and was slowing down the honeypot. Instead of Snort we decided to use Suricata, which does the same thing, but does it more efficiently.
+Unfortunately we noticed that Snort required a lot of resources and was slowing down the honeypot.
+Instead of Snort we decided to use Suricata, which uses the same ruleset, but uses it more efficiently and therefore less resources.
+
+\subsubsection{Dionaea}
+Dionaea is a low interaction honeypot that provides several services, namely http, ftp, tftp, MSSQL, MySQL and SIP (VoIP).
+We installed this to enable more services and attract more traffic to the honeypot.
 
@@
 
-\subsection{Auditing}
-For the HI phase, we would like to have a way to audit commandline interaction. 
-Various options exist, such as the user-space \verb|auditd| daemon, but to make auditing harder to detect this we feel it would be best to look for a kernel-space solution.
-An auditing module developed by the Honeynet project called Sebek can provide this, but is unmaintained and outdated.
-The grsec kernel patches also provide auditing from kernel space, but we were advised against replacing the stock kernel.
+\subsubsection{Auditing}
+During the High Interaction phase we used a kernel module that hooked execve (new process) calls and commandline interaction.
+This information was logged to syslog which provided us with auditing information during this phase.
+%For the HI phase, we would like to have a way to audit commandline interaction. 
+%Various options exist, such as the user-space \verb|auditd| daemon, but to make auditing harder to detect this we feel it would be best to look for a kernel-space solution.
+%An auditing module developed by the Honeynet project called Sebek can provide this, but is unmaintained and outdated.
+%The grsec kernel patches also provide auditing from kernel space, but we were advised against replacing the stock kernel.
 
-We are still investigating the best way to achieve useful auditing.
-A self-written kernel module that hooks system calls, albeit limited, might suffice to get some much needed auditing information from the honeypot.
+%We are still investigating the best way to achieve useful auditing.
+%A self-written kernel module that hooks system calls, albeit limited, might suffice to get some much needed auditing information from the honeypot.
 

### 2012-01-09 Paper_honeypot.tex
Attacks on services & discussion

@@
 
+\begin{itemize}
+\item Kippo
+
+The first service we will be talking about is Kippo. During the interaction phase we've had 2610 login attempt on our Kippo client and 3871 connections. 54 of these attempts were succesfull.
+We have made a top 5 of most tried usernames in these login attempts:
+
+\begin{itemize}
+\item   2206 root
+\item     39 admin
+\item     26 bin
+\item     20 w000t
+\item     19 oracle
+\item     19 jenny
+\end{itemize}
+
+It is not entirely unexpected that the username "root" is tried most ofted. 
+We assume the username "w000t" is a backdoor user created in one or more exploits.
+Interesting to note is that "jenny" is the name of a fake employee listed on the our honeypot website.
+Most usernames were tried multiple times, even if they made no sense to us (for instance msnayeem_iitd tried to log in four times).
+
+These are the top 5 passwords tried:
+
+\begin{itemize}
+\item     55 
+\item     28 123456
+\item     20 111111
+\item     16 root
+\item     16 admin
+\end{itemize}
+
+As you can see, the most tried password is a blank password.
+The other passwords in this top 5 are all standard password that can be found in a dictionary attack.
+It is nice to note that 123456 was accepted by our system.
+
+An interesting situation was someone that tried to log in to the honeypot with the user root, just trying a list of years (1960-1980).
+He obviously thought we would use a birthyear as our password (which was not the case).
+
+\item HTTP
+In three weeks, a total of 76002 HTTP-requests were received by our Apache server originating from 176 unique IP addresses. 4369 of these requests were 404'd.
+
+These 404's are interesting because they show us what adresses the attackers were trying to access.
+For instance, we see a lot of scans for phpMyAdmin (phpMyAdmin-3.3.10.5, phpMyAdmin-3.4.7, phpMyAdmin-3.4.6 etcetera).
+They also searched non existing folders for these phpMyAdmin files.
+
+Also serveral client IDs are visible in these logs: "\verbatim{"-"}", which is a blank user/client ID, (3432 attempts), "MiRRORS" (310 attempts), "ZmEu" (41 attempts) and "Morfeus Fucking Scanner" (4 attempts).
+These seem to be search engine crawlers.
+It is interesting to see that both several "\verbatim{"-"}" and several "ZmEu" users are scanning for "/w00tw00t.at.blackhats.romanian.anti-sec:)" and "/w00tw00t.at.ISC.SANS.DFind:)".
+The attackers using "ZmEu" found one valid path: /phpmyadmin/, however he did not exploit this.
+"Morfeus Fucking Scanner" was used by one attacker, that soly searched for /user/soapCaller.bs.
+We could not find anything on this .bs file.
+We expected to see several occurances of "/robots.txt" because this is a common way to start when using a search engine crawler, but none were found.
+
 \section{Connection between the attacks}
@@
 
+Our honeypot was not as popular as we had hoped it would be.
+Because we chose not to advertise, we solely relied on the hackers abilities to find us by for example portscans.
+The number of attacks on us was therefore not high and neither was the depth of the attacks.
+
+Concluding:
+
 \section{Discussion}
@@
 \item Low/High interaction phase
-\item GHH
-\item etc
+
+To see what the best approach is to a honeypot project, we decided to split the project in a Low and a High interaction phase.
+Unfortunately the High interaction phase yielded not much more results than the low interaction phase.
+Additionally we were a little late in switching to the HI phase.
+So in hindsight we don't think it was worth switching to the high interaction phase.
+
+\item Lack of interest
+
+We decided to make this honeypot as realistic as possible, this meant: not advertising our honeypot on IRC-channels or 4Chan.
+Our motivation was that in nonhoneypot circumstances this would also not be done.
+So for example advertising the IP address of the honeypot would attract a lot more hackers than a normal server would.
+So the results would not be on par with attacks against a normal server.
+In the High interaction phase we had a website that attracted people to our honeypot, which is a more realistic approach to attracting hackers.
+
+\item DenyHosts
+
+To make sure our real SSH port would not be brute forced, we added DenyHost to the port.
+We soon realized that this was a very effective countermeasurement versus intruders: both Bart and Joost were blocked due to their lack of Linux knowledge.
+Luckely Mark and Aram were able to unblock the other two.
+
+\item Linux knowledge
+
+Two of the teammembers had limited Linux experience.
+It took some time for them to understand how Linux commandline works.
+Fortunately we had two guru's in the group that were able to teach the newbies.
+
+\item Google Hack Honeypot
+
+The Google Hack Honeypot was supposed to attract hackers.
+Unfortunately it soon crashed due to unknown configuration errors.
+Because the results in the period the tool was running were nonexistent and we were switching to HI later on, we decided not to use it.
+
 \end{itemize}

### 2012-01-09 Paper_honeypot.tex
wat aanpassingen

@@
 They also searched non existing folders for these phpMyAdmin files.
-
 Also serveral client IDs are visible in these logs: "\verbatim{"-"}", which is a blank user/client ID, (3432 attempts), "MiRRORS" (310 attempts), "ZmEu" (41 attempts) and "Morfeus Fucking Scanner" (4 attempts).
@@
 The attackers using "ZmEu" found one valid path: /phpmyadmin/, however he did not exploit this.
-"Morfeus Fucking Scanner" was used by one attacker, that soly searched for /user/soapCaller.bs.
+Morfeus Fucking Scanner was used by one attacker, that soly searched for /user/soapCaller.bs.
 We could not find anything on this .bs file.
@@
 \label{Conclusion}
-From Section \ref{Connection} we can conclude that...
-
-Our honeypot was not as popular as we had hoped it would be.
+From Section \ref{Connection} we can conclude that our honeypot was not as popular as we had hoped it would be.
 Because we chose not to advertise, we solely relied on the hackers abilities to find us by for example portscans.
 The number of attacks on us was therefore not high and neither was the depth of the attacks.
-
-Concluding:
 
@@
 
+\begin{itemize}
+\item Low/High interaction phase
+\item GHH
+\item etc
+\end{itemize}
+
 %%% REFS %%%

### 2012-01-10 Paper_honeypot.tex
gefixed .tex compiled nu wel

--- 2012-01-09 Paper_honeypot.tex
@@
 
-\begin{itemize}
-\item Kippo
+\subsection{Kippo}
 
@@
 Interesting to note is that "jenny" is the name of a fake employee listed on the our honeypot website.
-Most usernames were tried multiple times, even if they made no sense to us (for instance msnayeem_iitd tried to log in four times).
+Most usernames were tried multiple times, even if they made no sense to us for instance msnayeem\_iitd tried to log in four times.
 
@@
 
-\item HTTP
+\subsection{HTTP}
 In three weeks, a total of 76002 HTTP-requests were received by our Apache server originating from 176 unique IP addresses. 4369 of these requests were 404'd.
@@
 \label{Conclusion}
-From Section \ref{Connection} we can conclude that...
+From Section \ref{Connection} we can conclude that our honeypot was not as popular as we had hoped it would be.
+Because we chose not to advertise, we solely relied on the hackers abilities to find us by for example portscans.
+The number of attacks on us was therefore not high and neither was the depth of the attacks.
 

### 2012-01-10 Paper_honeypot.tex
Eerste aanpassing layout & packages toegevoegd

@@
 
+\usepackage{booktabs}
+\usepackage[parfill]{parskip}
 %-------------------------------- Title ----------------------------------
@@
 
-\begin{itemize}
-\item   2206 root
-\item     39 admin
-\item     26 bin
-\item     20 w000t
-\item     19 oracle
-\item     19 jenny
-\end{itemize}
+\begin{tabular}{|l|r|}
+\hline
+2206 & root \\ \hline
+39 & admin \\ \hline
+26 & bin \\ \hline
+20 & w000t \\ \hline
+19 & oracle \\ \hline
+19 & jenny \\ \hline
+\end{tabular}
 
@@
 
-\begin{itemize}
-\item     55 
-\item     28 123456
-\item     20 111111
-\item     16 root
-\item     16 admin
-\end{itemize}
+\begin{tabular}{|l|r|}
+\hline
+55 &  \\ \hline
+28 & 123456 \\ \hline
+20 & 111111 \\ \hline
+16 & root \\ \hline
+16 & admin \\ \hline
+\end{tabular}
 
@@
 
-These 404's are interesting because they show us what adresses the attackers were trying to access.
+These 404's are interesting because they show us what addresses the attackers were trying to access.
 For instance, we see a lot of scans for phpMyAdmin (phpMyAdmin-3.3.10.5, phpMyAdmin-3.4.7, phpMyAdmin-3.4.6 etcetera).
@@
 
-Also serveral client IDs are visible in these logs: "\verbatim{"-"}", which is a blank user/client ID, (3432 attempts), "MiRRORS" (310 attempts), "ZmEu" (41 attempts) and "Morfeus Fucking Scanner" (4 attempts).
-These seem to be search engine crawlers.
-It is interesting to see that both several "\verbatim{"-"}" and several "ZmEu" users are scanning for "/w00tw00t.at.blackhats.romanian.anti-sec:)" and "/w00tw00t.at.ISC.SANS.DFind:)".
-The attackers using "ZmEu" found one valid path: /phpmyadmin/, however he did not exploit this.
-"Morfeus Fucking Scanner" was used by one attacker, that soly searched for /user/soapCaller.bs.
-We could not find anything on this .bs file.
-We expected to see several occurances of "/robots.txt" because this is a common way to start when using a search engine crawler, but none were found.
+Also serveral client IDs are visible in these logs:
+
+\begin{itemize}
+\item ``\emph{``-''}'' (3432 attempts)
+
+This is a blank user/client ID.
+
+\item ``MiRRORS'' (310 attempts)
+
+\tocheck{Info over MiRRORS}
+
+\item ``ZmEu'' (41 attempts)
+
+The attackers using ``ZmEu'' found one valid path: /phpmyadmin/, however he did not exploit this.
+It is interesting to see that both several ``\emph{''-''}'' and several ``ZmEu'' users are scanning for ``/w00tw00t.at.blackhats.romanian.anti-sec:)'' and ``/w00tw00t.at.ISC.SANS.DFind:)''
+
+\item ``Morfeus Fucking Scanner'' (4 attempts)
+
+``Morfeus Fucking Scanner'' was used by one attacker, that soly searched for /user/soapCaller.bs. We could not find anything on this .bs file.
+\end{itemize}
+
+We expected to see several occurances of ``/robots.txt'' because this is a common way to start when using a search engine crawler, but none were found.
 

### 2012-01-10 Paper_honeypot.tex
Verdere aanpassingen

@@
 
-The first service we will be talking about is Kippo. During the interaction phase we've had 2610 login attempt on our Kippo client and 3871 connections. 54 of these attempts were succesfull.
+The first service we will be talking about is Kippo. During the interaction phase we've had 2610 login attempt on our Kippo client and 3871 connections. 54 of these attempts were successful.
 We have made a top 5 of most tried usernames in these login attempts:
@@
 
-It is not entirely unexpected that the username "root" is tried most ofted. 
-We assume the username "w000t" is a backdoor user created in one or more exploits.
-Interesting to note is that "jenny" is the name of a fake employee listed on the our honeypot website.
-Most usernames were tried multiple times, even if they made no sense to us for instance msnayeem\_iitd tried to log in four times.
+It is not entirely unexpected that the username ``root'' is tried most often. 
+The username ``w000t'' is a backdoor user created in the FAKE Apache 2.2.17 exploit.
+Interesting to note is that ``jenny'' is the name of a fake employee listed on the our honeypot website.
+Most usernames were tried multiple times, even if they made no sense to us for instance ``msnayeem\_iitd'' tried to log in four times.
 
@@
 An interesting situation was someone that tried to log in to the honeypot with the user root, just trying a list of years (1960-1980).
-He obviously thought we would use a birthyear as our password (which was not the case).
+It seems like an unusual brute force attack.
+Assuming someone used their birthday as a password, which still happens a lot, it is unlikely that this someone just used their birth year and not the whole birthday.
+However brute forcing for the entire birthday takes way to long.
+So for a quick and easy brute force it is unlikely to yield results.
 
@@
 
-Also serveral client IDs are visible in these logs:
+Also several client IDs are visible in these logs:
 
 \begin{itemize}
-\item ``\emph{``-''}'' (3432 attempts)
+\item \emph{``-''} (3432 attempts)
 
-This is a blank user/client ID.
+This is a blank user/client ID. Which seems to be a popular client to use when you want to find weak software on a server.
 
@@
 
-\tocheck{Info over MiRRORS}
+Due to the well-chosen name we could not find additional information on this client. Mirrors is a frequently used word on the internet.
 
@@
 The attackers using ``ZmEu'' found one valid path: /phpmyadmin/, however he did not exploit this. 
-It is interesting to see that both several ``\emph{''-''}'' and several ``ZmEu'' users are scanning for ``/w00tw00t.at.blackhats.romanian.anti-sec:)'' and ``/w00tw00t.at.ISC.SANS.DFind:)''
+It is interesting to see that both several \emph{''-''} and several ``ZmEu'' users are scanning for ``/w00tw00t.at.blackhats.romanian.anti-sec:)'' and ``/w00tw00t.at.ISC.SANS.DFind:)''
 
@@
 
-``Morfeus Fucking Scanner'' was used by one attacker, that soly searched for /user/soapCaller.bs. We could not find anything on this .bs file.
+``Morfeus Fucking Scanner'' was used by one attacker, that solely searched for /user/soapCaller.bs. We could not find anything on this .bs file.
 \end{itemize}
 
-We expected to see several occurances of ``/robots.txt'' because this is a common way to start when using a search engine crawler, but none were found.
+We expected to see several occurrences of ``/robots.txt'' because this is a common way to start when using a search engine crawler, but none were found.
 
@@
 
-\begin{itemize}
-\item Low/High interaction phase
-\item GHH
-\item etc
-\end{itemize}
-
 %%% REFS %%%

### 2012-01-10 Paper_honeypot.tex
Uitbreiding Attacks, Tables omgedraaid.

@@
 \hline
-2206 & root \\ \hline
-39 & admin \\ \hline
-26 & bin \\ \hline
-20 & w000t \\ \hline
-19 & oracle \\ \hline
-19 & jenny \\ \hline
+Username & Attempts \\ \hline
+root & 2206 \\ \hline
+admin & 39 \\ \hline
+bin & 26 \\ \hline
+w000t & 20 \\ \hline
+oracle & 19 \\ \hline
+jenny & 19 \\ \hline
 \end{tabular}
@@
 \hline
-55 &  \\ \hline
-28 & 123456 \\ \hline
-20 & 111111 \\ \hline
-16 & root \\ \hline
-16 & admin \\ \hline
+Password & Attempts \\ \hline
+  & 55 \\ \hline
+123456 & 28 \\ \hline
+111111 & 20 \\ \hline
+root & 16 \\ \hline
+admin & 16 \\ \hline
 \end{tabular}
@@
 \label{Conclusion}
-From Section \ref{Connection} we can conclude that our honeypot was not as popular as we had hoped it would be.
-Because we chose not to advertise, we solely relied on the hackers abilities to find us by for example portscans.
-The number of attacks on us was therefore not high and neither was the depth of the attacks.
-
-Our honeypot was not as popular as we had hoped it would be.
+From Section \ref{Attacks} we can conclude that our honeypot was not as popular as we had hoped it would be.
 Because we chose not to advertise, we solely relied on the hackers abilities to find us by for example portscans.

### 2012-01-10 Paper_honeypot.tex
added some info about the dionaea logs

@@
 
+\subsection{Dionaea}
+As mentioned in Section \ref{Setup} Dionaea runs several services. In the table below we will show how many times those services got connected.
+
+\begin{tabular}{|l|c|r|}
+\hline
+Service & Port & Number of connections \\ \hline
+MSSQL  & 1433 & 1632 \\ \hline
+MS SMB filesharing & 445 & 67 \\ \hline
+FTP & 21 & 55 \\ \hline
+MySQL & 3306 & 54 \\ \hline
+Session Initiation Protocol & 5060 & 9 \\ \hline
+\end{tabular}
+
+As we looked further into the Dionaea logfiles we found that MSSQL was connected to from 95.82.11.28, which is in Iran, 1414 times.
+However, the SMB filesharing, FTP and SIP were mostly connected from one IP, namely 130.89.165.58.
+This IP is from Twente University, so probably it was another student.
+
 \section{Connection between the attacks}

### 2012-01-10 Paper_honeypot.tex
Added iptables graphs

--- 2012-01-10 Paper_honeypot.tex
@@
 In this section we will discuss the statistical data that have been gathered from our logfiles. 
+The logged connection attempts have been processed to group them by the IP address' originating country through the use of WHOIS directory service information.
+It is worthwhile noting that a significant number of connecting hosts could not be referenced to RWHOIS information.
+Also note that these graphs display all logged connection attempts rather than successful connections, giving us a general view of which ports attract attention.
+
+For clarity we have chosen to graph only the 10 most active originating countries, and make use of a logarithmic scale to represent the spectrum of port numbers.
+The labels on the \(y\)-axis were chosen based on the most frequently accessed ports, with some selected labels removed to avoid clutter.
+
+\includegraphics[width=1.7\textwidth,angle=-90]{tcp}
+
+\includegraphics[width=1.7\textwidth,angle=-90]{udp}
 

### 2012-01-10 Paper_honeypot.tex
Connecties between the attacks uitgewerkt

@@
 The attackers using ``ZmEu'' found one valid path: /phpmyadmin/, however he did not exploit this. 
-It is interesting to see that both several \emph{''-''} and several ``ZmEu'' users are scanning for ``/w00tw00t.at.blackhats.romanian.anti-sec:)'' and ``/w00tw00t.at.ISC.SANS.DFind:)''
+It is interesting to see that both several \emph{''-''} and several ``ZmEu'' users are scanning for ``/w00tw00t.at.blackhats.romanian.anti-sec:)'' and ``/w00tw00t.at.ISC.SANS.DFind:)'' These clients are alleged to be for reconnacance purposes.
 
@@
 
-``Morfeus Fucking Scanner'' was used by one attacker, that solely searched for /user/soapCaller.bs. We could not find anything on this .bs file.
+``Morfeus Fucking Scanner'' was used by one attacker, that solely searched for /user/soapCaller.bs. Multiple sources say this might be a Drupal exploit.
 \end{itemize}
@@
 We will try to look for a pattern or similarities in this data and determine if the attack was more sophisticated.
+
+Three clients connected to both HTTP and Kippo, which raised our suspision:
+\begin{itemize}
+\item 58.17.52.14
+
+This citizen of the Peoples Republic of China poked every possible entrance, in two consequental days: he made a connection to Kippo and disconnected immideately.
+He also requested the main / folder via HTTP, using the client "-".
+Scanlogd also reported scans on port 80, 8080, 3128, 9415, 8909, 9592, 1080, 443.
+Unfortunately for us we were not attractive for him to return and actually break in.
+
+\item 61.190.172.2
+
+On four occasions this inhabitant of China probed us, split over 4 weeks.
+He also just connected to Kippo (Suricata reported a SSH-scan) and disconnected: he did this for fourty times in 3 minutes.
+Interestingly enough he was one of the users of the Morfius Fucking Scanner client: only requesting the /user/soapCaller.bs file.
+
+\item 125.88.75.199
+
+The last prober was also from China.
+On November 22th he connected to Kippo to automaticly disconnect again.
+He also requested the / folder via HTTP with the "-" client.
+He was only seen at this occasion.
+
+\end{itemize}
 

### 2012-01-10 Paper_honeypot.tex
Added PNG versions of graphs
RDNS comment
Spell Check

@@
 In the fourth section we will discuss which attacks took place against the different services the honeypot was running e.g. SSH and MySQL.
-In order to discover the sophisticationof the attacks, we will try to present possible connections between an attack on one service and an attack on a different service, in the fifth section.
+In order to discover the sophistication of the attacks, we will try to present possible connections between an attack on one service and an attack on a different service, in the fifth section.
 The sixth section will conclude our findings. 
@@
 On the 1st of November we delivered our initial setup document, however during the "Waiting for hackers" phase we have made adjustments to the setup in order to attract more activity to the honeypot.
-In this section we will present the initial setup and in addittion to that, discuss the changes that were made.
+In this section we will present the initial setup and in addition to that, discuss the changes that were made.
 
@@
 
-We have put the command to create the SSH tunnel in a script to allow the connection to be re-esthablished quickly and easily if the SSH session is terminated for some reason. 
+We have put the command to create the SSH tunnel in a script to allow the connection to be re-established quickly and easily if the SSH session is terminated for some reason. 
 A severance of the connection would lead us to assume the worst initially.
@@
 After collecting our syslog information, we wanted to be somehow alerted of interesting information automatically.
-We decided to use an IRC bot that would leave a message in an IRC channel (\#contagiarium) everytime an entry was made to the syslog file.
+We decided to use an IRC bot that would leave a message in an IRC channel (\#contagiarium) every time an entry was made to the syslog file.
 First this lead to an overloaded channel and several false positives.
@@
 To capture attackers that try to come in by brute-forcing SSH logins, we used Kippo. 
-Kippo is a sandbox SSH service, used to log bruteforce attacks and the shell interaction performed by the attacker after gaining access (with a trivial password). 
+Kippo is a sandbox SSH service, used to log brute force attacks and the shell interaction performed by the attacker after gaining access (with a trivial password). 
 The latest development version was installed to run as a separate user.
@@
 
-During the second phase of the project, we noticed that of the succesfull bruteforce attempts against Kippo, none actually tried to continue and explore or use the honeypot.
+During the second phase of the project, we noticed that of the successful brute force attempts against Kippo, none actually tried to continue and explore or use the honeypot.
 Running a default portscan with Metasploit revealed that it can detect if Kippo is being used.
@@
 If someone found out that our real SSH port was changed to 9022, they might try to break in through the real SSH service.
-To make sure no bruteforce attacks could be done on this port, DenyHosts was installed.
+To make sure no brute force attacks could be done on this port, DenyHosts was installed.
 This service blocks IP-addresses that try to log in more than 3 times with the wrong password by adding them to the \verb|hosts.deny| file.
@@
 These attackers will use search queries to find sensitive data, for example password files, that due to some misconfiguration ended up being indexed by a search engine like Google.
-GHH allows logging of these requests to a MySQL database or Comma Seperated Value (CSV) file.
+GHH allows logging of these requests to a MySQL database or Comma Separated Value (CSV) file.
 
@@
 
-We hoped to detect and capture a bruteforce attack or scan on generic filenames, for example \verb|passwd.txt|, because this would probably have warned us that an attacker will try to use a dictionary attack.
+We hoped to detect and capture a brute force attack or scan on generic filenames, for example \verb|passwd.txt|, because this would probably have warned us that an attacker will try to use a dictionary attack.
 Unfortunately GGH broke down after the Virtual Machines were shut down, we were unable to get it back up running and therefore decided not invest more time and effort into it.
@@
 To make it more appealing in ways of computing power we used the department of Partical Physics and Astronomy as a cover.
-On the website we also put some information about a fictive person (Jenny Abott) would manage the servers and website.
-So we created a useraccount with the same login and password, namely Jenny.
+On the website we also put some information about a fictive person (Jenny Abbott) would manage the servers and website.
+So we created a user account with the same login and password, namely Jenny.
 This way a more sophisticated attack would be possible.
-An attacker would look at the website, determine if it is intresing to attack and then connect to Kippo to log in as Jenny.
+An attacker would look at the website, determine if it is interesting to attack and then connect to Kippo to log in as Jenny.
 
 \subsubsection{Wordpress}
-In order to make the website more vunerable and appealing to hackers, we installed an out of date version of Wordpress.
+In order to make the website more vulnerable and appealing to hackers, we installed an out of date version of Wordpress.
 Wordpress is software, powered by PHP and MySQL, that can be used to easily create websites or blogs and manage them via the content management system.
@@
 
-We set up Snor using the community rule-set that will only alert on suspicious network behaviour, but do nothing to prevent it. 
+We set up Snort using the community rule-set that will only alert on suspicious network behaviour, but do nothing to prevent it. 
 Included are rules on detecting exploit attempts on various common services like HTTP, FTP, SMTP, and many more.
@@
 
-\includegraphics[width=1.7\textwidth,angle=-90]{tcp}
+\includegraphics[width=1.7\textwidth,angle=-90]{tcp.png}
 
-\includegraphics[width=1.7\textwidth,angle=-90]{udp}
+\includegraphics[width=1.7\textwidth,angle=-90]{udp.png}
 
@@
 The attackers using ``ZmEu'' found one valid path: /phpmyadmin/, however he did not exploit this. 
-It is interesting to see that both several \emph{''-''} and several ``ZmEu'' users are scanning for ``/w00tw00t.at.blackhats.romanian.anti-sec:)'' and ``/w00tw00t.at.ISC.SANS.DFind:)'' These clients are alleged to be for reconnacance purposes.
+It is interesting to see that both several \emph{''-''} and several ``ZmEu'' users are scanning for ``/w00tw00t.at.blackhats.Romanian.anti-sec:)'' and ``/w00tw00t.at.ISC.SANS.DFind:)'' These clients are alleged to be for reconnaissance purposes.
 
@@
 
-Three clients connected to both HTTP and Kippo, which raised our suspision:
+Three clients connected to both HTTP and Kippo, which raised our suspicion:
 \begin{itemize}
@@
 
-This citizen of the Peoples Republic of China poked every possible entrance, in two consequental days: he made a connection to Kippo and disconnected immideately.
+This citizen of the Peoples Republic of China poked every possible entrance, in two consequential days: he made a connection to Kippo and disconnected immediately.
 He also requested the main / folder via HTTP, using the client "-".
@@
 On four occasions this inhabitant of China probed us, split over 4 weeks.
-He also just connected to Kippo (Suricata reported a SSH-scan) and disconnected: he did this for fourty times in 3 minutes.
-Interestingly enough he was one of the users of the Morfius Fucking Scanner client: only requesting the /user/soapCaller.bs file.
+He also just connected to Kippo (Suricata reported a SSH-scan) and disconnected: he did this for forty times in 3 minutes.
+Interestingly enough he was one of the users of the Morfeus Fucking Scanner client: only requesting the /user/soapCaller.bs file.
 
@@
 \label{Discussion}
-In this section we would like to discuss some issues we encountered during the honeypotproject.
+In this section we would like to discuss some issues we encountered during the honeypot project.
 
@@
 We decided to make this honeypot as realistic as possible, this meant: not advertising our honeypot on IRC-channels or 4Chan.
-Our motivation was that in nonhoneypot circumstances this would also not be done.
+Our motivation was that in non-honeypot circumstances this would also not be done.
 So for example advertising the IP address of the honeypot would attract a lot more hackers than a normal server would.
@@
 To make sure our real SSH port would not be brute forced, we added DenyHost to the port.
-We soon realized that this was a very effective countermeasurement versus intruders: both Bart and Joost were blocked due to their lack of Linux knowledge.
-Luckely Mark and Aram were able to unblock the other two.
+We soon realized that this was a very effective counter measurement versus intruders: both Bart and Joost were blocked due to their lack of Linux knowledge.
+Luckily Mark and Aram were able to unblock the other two.
 
@@
 
-Two of the teammembers had limited Linux experience.
+Two of the team members had limited Linux experience.
 It took some time for them to understand how Linux commandline works.
@@
 Unfortunately it soon crashed due to unknown configuration errors.
-Because the results in the period the tool was running were nonexistent and we were switching to HI later on, we decided not to use it.
+Because the results in the period the tool was running were non-existent and we were switching to HI later on, we decided not to use it.
 

### 2012-01-10 Paper_honeypot.tex
Conclusion

@@
 It is worthwhile noting that a significant number of connecting hosts could not be referenced to RWHOIS information.
+In these cases we have attempted to resolve the missing information by making use of reverse DNS queries and grouping by associated top level domain.
 Also note that these graphs display all logged connection attempts rather than successful connections, giving us a general view of which ports attract attention.
@@
 \label{Conclusion}
-From Section \ref{Attacks} we can conclude that our honeypot was not as popular as we had hoped it would be.
-Because we chose not to advertise, we solely relied on the hackers abilities to find us by for example portscans.
-The number of attacks on us was therefore not high and neither was the depth of the attacks.
+The Setup we used for this honeypot seems to be sufficient.
+In both phases of the research we did not lose control over our honeypot, which was our main worry.
+We also decided not to advertise, which gave our honeypot less traffic, but a more realistic outcome.
 
-Concluding:
+From the Section Attacks we can conclude that our honeypot was not as popular as we had hoped it would be.
+We blame this on not advertising our IP.
+The numbers of attacks on us were therefore not high and neither was the depth of the attacks.
+
+Only three real scans took place, as reported in the Section Connections between attacks. Unfortuntaly we did not make our honeypot interesting enough for them to return and actually break in.
+
+Our chose for swiching from a Low Interaction to a High Interaction honeypot seems like a good tactic, but due to downtime and bugged software, we did not have enough time.
+
+Concluding: we did see some people checking us out, but noone actually broke in. For further research we suggest using the same setup, but having the server online for a much longer time.
 

### 2012-01-10 Paper_honeypot.tex
Reviewing setup

--- 2012-01-10 Paper_honeypot.tex
@@
 \title{\textbf{Honeypot Assignment: \\ \emph{Group Contagiarium}}}
-\author{Aram Verstegen, \tocheck{studentnummer Utwente} \\
+\author{Aram Verstegen, s4092368 \\
 	 Mark Vijfvinkel, 1275194 \\
@@
 
-\subsection{Monitoring Machine}
-
-In this subsection we describe the setup of the monitoring machine.
-
-\subsubsection{Syslog-ng}
-By default, \verb|syslog| uses UDP to transmit its log messages to a remote server.
-Relying on UDP is not desirable, because everything is sent in plaintext and delivery of UDP messages is not guaranteed.
-To allow for confidential and reliable transmission of the syslog messages from the one machine to another, one needs to use TCP connections in syslog, and provide some way of encapsulating the socket used in a cryptographic layer.
-
-We decided to set up a reverse SSH tunnel from the monitoring machine to the honeypot machine, as described in the article "Remote logging with SSH and syslog-ng''. \cite{remote_logging}
-The example command establishes an SSH connection the honeypot, setting options to forego a login shell or anything a normal user would desire, and binds the local socket 9514 on the honeypot machine to tunnel its connections to the same port on our monitoring machine.
-
-We configured a locally installed \verb|syslog-ng| package for our user on the monitoring machine and have configured it to listen to local TCP connections on port 9514 and will thereby receive syslog messages transmitted through this port.
-%We are aware there is still the possibility of other users on the monitoring machine poisoning our logs, but we feel we can rule out foul play by our peers in this assignment.
-
-To allow for secure SSH access, a public keypair was generated on the util machine and the associated public key was added to the honeypot's \verb|authorized_keys| file.
-
-We have put the command to create the SSH tunnel in a script to allow the connection to be re-esthablished quickly and easily if the SSH session is terminated for some reason. 
-A severance of the connection would lead us to assume the worst initially.
-If re-establishing this connection fails after any attackers have gained access (which should never happen if we monitor adequately), we would have assumed that the machine was compromised beyond our control and would have requested a shutdown of the VM, if no network downtime was otherwise announced.
-
-\subsubsection{Monitoring}
-After collecting our syslog information, we wanted to be somehow alerted of interesting information automatically.
-We decided to use an IRC bot that would leave a message in an IRC channel (\#contagiarium) everytime an entry was made to the syslog file.
-First this lead to an overloaded channel and several false positives.
-To solve these problems filters (regular expressions in a file) were created to only post relevant information.
-For example, from the kippo logs only the actual login attempts would be shown with the tried password.
-
-\subsection{Honeypot}
-
-In this section we describe the setup of the honeypot.
-
-\subsubsection{Rsyslog}
-On the honeypot, we modified the default \verb|rsyslog| configuration to also send syslog messages to the (now bound) local port 9514, completing the chain to the syslog instance on our monitoring machine.
-At first the modified configuration was stored in its own file in the syslog.d directory, which is a bit obvious to anyone taking an interest.
-Therefore we moved the configuration to the standard syslog directory.
-
-%We plan to move this small configuration over to one of the default configuration files to avoid raising suspicion outright.
-Note that the chosen port for remote syslog is very similar to the default syslog port 514, which might have given us away. So we changed the port number to 9514.
-%We plan to change it to a more nondescript port number.
-
-\subsubsection{SSHFS mount}
-Not all services support logging to syslog and for some it is infeasible.
-We created a read-only \verb|sshfs| mount from our monitoring machine that we used to write our logfiles for various services to.
-We used an interactive login to mount this, as allowing the honeypot back into our monitoring machine with an SSH keypair just seems dangerous.
-
-\subsubsection{scanlogd}
-To detect port scans we installed the service \verb|scanlogd|.
-Per detected scan an entry is written to syslog.
-
-\subsubsection{iptables}
-To log all incoming and outgoing connection attempts, we have added the following rules to the iptables configuration. The attempts were also logged to syslog.
-
-\begin{verbatim}
-iptables -I INPUT -m state --state NEW -j LOG --log-prefix "New Connection: "
-iptables -I OUTPUT -m state --state NEW -j LOG --log-prefix "New Connection: "
-\end{verbatim}
-
-We have filtered outgoing SMTP, SIP and NetBIOS traffic, in order to prevent abuse and spam.
-%We are also going to filter outgoing SMTP, SIP and NetBIOS traffic.
+\subsection{Rsyslog}
+At first the rsyslog configuration was stored in its own file in the syslog.d directory, which might have been seen by an attacker.
+Therefore we have moved it to the default syslog configuration file.
 
 \subsubsection{Kippo}
-To capture attackers that try to come in by brute-forcing SSH logins, we used Kippo. 
-Kippo is a sandbox SSH service, used to log bruteforce attacks and the shell interaction performed by the attacker after gaining access (with a trivial password). 
-The latest development version was installed to run as a separate user.
-
-To make this as realistic as possible we added a redirect in the iptables configuration from the real SSH port 22 to port 2222 that Kippo uses by default.
-The reason for forwarding in iptables was to not give Kippo root rights, which would be very dangerous.
-
-%Unfortunately the forward broke down around the time the firewall was being reconfigured, for reasons yet unclear . 
-%If it can't be resolved, we can circumvent this issue by running Kippo through a copy of the Python interpreter that has the capability to bind privileged ports.
-
-To be able to still log on to the honeypot, we configured our real SSH service to run on port 9022.
-
-Furthermore, we patched Kippo to log messages to syslog.
-
-During the second phase of the project, we noticed that of the succesfull bruteforce attempts against Kippo, none actually tried to continue and explore or use the honeypot.
-Running a default portscan with Metasploit revealed that it can detect if Kippo is being used.
-We thought this explained why nobody tried to log in.
-To solve this problem, we adjusted the Kippo configuration so that Metasploit could not recognize Kippo.
-Unfortunately this did still not lead to actual exploration or use of the honeypot.
+During the second phase of the project, we noticed that of the succesful dictionary-based bruteforce attempts against Kippo, none actually tried to continue and explore or use the honeypot.
+Running an nmap portscan from Metasploit (which adds to the information provided by nmap) revealed that it can detect if Kippo is being used.
+This doesn't explain why people should refrain from logging in after a successful dictionary attack
 

### 2012-01-10 Paper_honeypot.tex

@@
 During the second phase of the project, we noticed that of the succesful dictionary-based bruteforce attempts against Kippo, none actually tried to continue and explore or use the honeypot.
-Running an nmap portscan from Metasploit (which adds to the information provided by nmap) revealed that it can detect if Kippo is being used.
-This doesn't explain why people should refrain from logging in after a successful dictionary attack
+Running an \verb|nmap| portscan from the Metasploit framework (which adds to the information provided by \verb|nmap|) revealed that it can be detected if Kippo (or rather Python's twisted framework) is being used.
+We learned that twisted can be identified because it sends the supported ciphers during the key exchange phase before it has been prompted to do so.
+Still, this doesn't explain why people should refrain from logging in after a successful dictionary attack, as a sophisticated attacker would not pursue a dictionary attack if he were aware of the service acting as a honeypot.
 
-\subsubsection{DenyHosts}
-If someone found out that our real SSH port was changed to 9022, they might try to break in through the real SSH service.
-To make sure no bruteforce attacks could be done on this port, DenyHosts was installed.
-This service blocks IP-addresses that try to log in more than 3 times with the wrong password by adding them to the \verb|hosts.deny| file.
-The hosts listed there are blacklisted on the machine, but we can edit them if needed.
-
-%Denyhosts can log to syslog by editing its configuration file.
-
-\subsection{Apache \& Nginx}
-To attract web traffic, we installed the Apache web server, the Apache PHP runtime and the \verb|mod-security| module which allows us to spoof its version information, to make it appear old and vulnerable.
-The Nginx webserver is installed as a reverse proxy in front of Apache, which runs on local port 8080.
-This provides 2 layers of logging (to files), and Nginx's superior transparent caching mechanism allows us to avoid exhausting our resources with scripting languages, should we have ever faced an application-layer DoS attack (Nginx appears quite resilient against those).
-
-Nginx has been compiled from source and was patched to send the spoofed Apache version header as well.
-We will configure both services to log to our \verb|sshfs| mount.
-
-\subsubsection{Google Hack Honeypot}
-Google Hack Honeypot (GHH) is designed to be used to attract search engine attackers. 
-These attackers will use search queries to find sensitive data, for example password files, that due to some misconfiguration ended up being indexed by a search engine like Google.
-GHH allows logging of these requests to a MySQL database or Comma Seperated Value (CSV) file.
-
-Since GHH offers multiple search targets, we picked one to start with, namely GGH version 1.2 passlist.txt.
-We had no particular reason for this, only that it was easy to set up. 
-%We plan to install more variants of the GHH to attract more attackers, but before this could happen the VM's were shut down.
-
-%To set up GHH we have first installed Apache and PHP5.
-We configured GHH to log to in CSV format, since it requires minimal setup and is easily parsed.
-%We can later reconfigure to use MySQL, if the results are not satisfactory.
-Then, we patched the GHH code to allow logging to syslog, to maintain a single, consistent point of log collection.
-
-We hoped to detect and capture a bruteforce attack or scan on generic filenames, for example \verb|passwd.txt|, because this would probably have warned us that an attacker will try to use a dictionary attack.
-Unfortunately GGH broke down after the Virtual Machines were shut down, we were unable to get it back up running and therefore decided not invest more time and effort into it.
-We also believe it would not have brought us more interesting data.
+\subsubsection{Google hack honeypot}
+We hoped to detect and capture a brute force attack or scan on generic filenames, for example \verb|passwd.txt|, because this would probably have warned us that an attacker will try to use a dictionary attack.
+In retrospect, given that our honeypot does not provide an actual service, it is easy to identify dictionary-based scans in our logs.
+Unfortunately our configured GGH broke after an upgrade, and we decided not to spend more time on configuring it.
+Through other teams we became aware of the HIHAT software package, which might be a more useful alternative.
+%TODO references?
 
@@
 To attract more activity we decided to create a fake website that would resemble a website of the University of Twente.
-To make it more appealing in ways of computing power we used the department of Partical Physics and Astronomy as a cover.
-On the website we also put some information about a fictive person (Jenny Abott) would manage the servers and website.
-So we created a useraccount with the same login and password, namely Jenny.
+To make it more appealing in ways of computing power we assumed the fictional identity of ``The department of Partical Physics and Astronomy of the University of Twente'' as a cover.
+On the website we also put some information about a fictive person (Jenny Abbott) would manage the servers and website.
+So we created a user account with the same login and password, namely Jenny.
 This way a more sophisticated attack would be possible.
-An attacker would look at the website, determine if it is intresing to attack and then connect to Kippo to log in as Jenny.
+An attacker would look at the website, determine if it is interesting to attack and then connect to Kippo to log in as Jenny.
 
 \subsubsection{Wordpress}
-In order to make the website more vunerable and appealing to hackers, we installed an out of date version of Wordpress.
+In order to make the website more vulnerable and appealing to hackers, we installed an out of date version of Wordpress.
 Wordpress is software, powered by PHP and MySQL, that can be used to easily create websites or blogs and manage them via the content management system.
@@
 \subsubsection{Snort and Suricata}
-Snort is an open source intrusion detection system (IDS) that uses realtime traffic analysis and packet logging on internet protocol networks.
-Snort was designed to be used as a network IDS, however we will be running it on our host machine, effectively making it a host-based IDS.
-
-We set up Snor using the community rule-set that will only alert on suspicious network behaviour, but do nothing to prevent it. 
-Included are rules on detecting exploit attempts on various common services like HTTP, FTP, SMTP, and many more.
-
-We configured Snort to send alerts to syslog, and to log its packet captures to the read-only \verb|sshfs| mounted from our monitoring machine.
-We also installed \verb|oinkmaster| to automatically retrieve the latest community Snort rules, with a Snort API key registered specifically for this project.
-
-Unfortunately we noticed that Snort required a lot of resources and was slowing down the honeypot.
-Instead of Snort we decided to use Suricata, which uses the same ruleset, but uses it more efficiently and therefore less resources.
+Unfortunately we noticed that snort required a lot of resources and was slowing down the honeypot. We replaced snort with suricata, which is compatible with the snort ruleset format.
+This also allowed us to keep up with the snort community rules, which were refreshed daily using \verb|oinkmaster|.
+Furthermore, Suricata allows use of more advanced packet capture interfaces than \verb|libpcap|.
+We've configured suricata to sniff packets using the PF\_RING capture interface, a kernel-mode ring buffer of packets exposed via \verb|procfs|.
+This provided great results and allowed us to use many rules on a limited amount of memory.
 
 \subsubsection{Dionaea}
-Dionaea is a low interaction honeypot that provides several services, namely http, ftp, tftp, MSSQL, MySQL and SIP (VoIP).
-We installed this to enable more services and attract more traffic to the honeypot.
+Dionaea is a low interaction honeypot daemon that provides several services, namely HTTP, FTP, TFTP, MSSQL, MySQL and SIP (VoIP).
+We have installed this to expose more services after noticing failing probes on some of the associated ports of FTP, MySQL, MSSQL and SIP services. 
+Enabling MSSQL on a Linux server might raise suspicions, however we felt this single exception to the rule might be overlooked or otherwise forgiven as a normal NAT configuration.
+The number of connection requests to this port was so high we decided we had to see what attackers were looking for.
 
@@
 \subsubsection{Auditing}
-During the High Interaction phase we used a kernel module that hooked execve (new process) calls and commandline interaction.
-This information was logged to syslog which provided us with auditing information during this phase.
+We reviewed various solutions to audit the command line after a successful break-in in our setup document, none of which seemed to match exactly what we needed.
+The problem can be characterized as: ``From `stealthy', `maintained', and `works out of the box', choose two.''
+Finally, after much research we produced a simple self-hiding kernel module that hooks \verb|execve| (new process) calls and logs the entire argument string for each call to syslog.
+The module makes use of the \verb|jprobes| interface for kernel debugging, which turned out to be ideal for this task.
+After it has been hidden using a `magic' command sequence, the module cannot be detected in \verb|lsmod| output, and is not detectable by \verb|rkhunter| or \verb|chkrootkit|, nor can its symbols be seen in kernel symbol lists exported by \verb|procfs|.
+%TODO reference phrack
+
 %For the HI phase, we would like to have a way to audit commandline interaction. 
@@
 The logged connection attempts have been processed to group them by the IP address' originating country through the use of WHOIS directory service information.
-It is worthwhile noting that a significant number of connecting hosts could not be referenced to RWHOIS information.
+It is worthwhile noting that a significant number of connecting hosts could not be referenced to WHOIS information.
 In these cases we have attempted to resolve the missing information by making use of reverse DNS queries and grouping by associated top level domain.
@@
 On four occasions this inhabitant of China probed us, split over 4 weeks.
-He also just connected to Kippo (Suricata reported a SSH-scan) and disconnected: he did this for fourty times in 3 minutes.
-Interestingly enough he was one of the users of the Morfius Fucking Scanner client: only requesting the /user/soapCaller.bs file.
+He also just connected to Kippo (suricata reported a SSH-scan) and disconnected: he did this for forty times in 3 minutes.
+Interestingly enough he was one of the users of the Morfeus Fucking Scanner client: only requesting the /user/soapCaller.bs file.
 

### 2012-01-10 Paper_honeypot.tex
Conclusion edit

@@
 
-Only three real scans took place, as reported in the Section Connections between attacks. Unfortuntaly we did not make our honeypot interesting enough for them to return and actually break in.
+Only three real scans took place, as reported in the Section Connections between attacks. Unfortunately we did not make our honeypot interesting enough for them to return and actually break in.
 
-Our chose for swiching from a Low Interaction to a High Interaction honeypot seems like a good tactic, but due to downtime and bugged software, we did not have enough time.
+Our choice for swiching from a Low Interaction to a High Interaction honeypot seems like a good tactic, but due to the low activity we did not have enough time to collect significant results.
 
-Concluding: we did see some people checking us out, but noone actually broke in. For further research we suggest using the same setup, but having the server online for a much longer time.
+Concluding: we did see a lot of probing activity, but no one actually broke in.
+%For further research we suggest using the same setup, but having the server online for a much longer time.
+We would like to continue a similar setup for research purposes privately.
 
@@
 Two of the team members had limited Linux experience.
-It took some time for them to understand how Linux commandline works.
+It took some time for them to understand how Linux command line works.
 Fortunately we had two guru's in the group that were able to teach the newbies.
@@
 
-The Google Hack Honeypot was supposed to attract hackers.
+The Google Hack Honeypot was supposed to entice hackers.
 Unfortunately it soon crashed due to unknown configuration errors.
 Because the results in the period the tool was running were non-existent and we were switching to HI later on, we decided not to use it.
+
+\item Lack of tools
+
+There exists a significant lack of tools to stealthily audit HI Linux honeypot activity, and as can by Project Honeynet's Google summer of Code project, a lack of graphing tools for network data.
+For this paper we invented our own using Python and Gnuplot.
 

### 2012-01-13 honeypot_software/iptables_log_parser.py
Added gnuplot/python script for iptables logs

@@ -1,0 +1,127 @@
+from datetime import datetime
+import re
+import subprocess
+import pickle
+import os
+import time
+import sys
+import operator
+
+whois_db = {}
+top_ports = {'tcp': {}, 'udp': {}}
+WHOIS_FILE = 'whois_cache'
+
+def iptables_syslog_line_to_datetime_ip_tuple(line):
+  start = line[0:15]
+  dt = datetime.strptime("%s 2011" % start, "%b %d %H:%M:%S %Y")
+  ip = re.search('SRC=([^ ]+)', line).groups()[0]
+  dpt = re.search('DPT=([^ ]+)', line).groups()[0]
+  country = whois_lookup(ip)
+  return (dt, ip, dpt, country)
+
+def whois_lookup(ip):
+  if ip not in whois_db:
+    try:
+      data = subprocess.check_output(['whois', ip])
+      whois_db[ip] = re.search('country:\s(.*)', str(data), flags=re.IGNORECASE).groups(0)[0].strip().upper()
+    except Exception, e:
+      try:
+        data = subprocess.check_output(['host', ip])
+        whois_db[ip] = data.split('.')[-2].upper()
+      except Exception, e:
+        whois_db[ip] = '?'
+  return whois_db[ip]
+
+def parse_file(filename):
+  logs_per_country = {'udp': {}, 'tcp': {}}
+  loglines = file(filename).readlines()
+
+  for line in loglines:
+    proto = None
+    if 'New Connection: IN=eth1' in line and not 'SRC=130.89.145.99' in line and not 'PROTO=ICMP' in line:
+      entry = line.strip()
+      if 'PROTO=TCP' in line:
+        proto = 'tcp'
+      if 'PROTO=UDP' in line:
+        proto = 'udp'
+
+      if not proto:
+        continue
+
+      try:
+        (timestamp, ip, port, country) = iptables_syslog_line_to_datetime_ip_tuple(entry)
+      except Exception, e:
+        print "Error %s" % e
+        print entry
+
+      if not country in logs_per_country[proto]:
+        logs_per_country[proto][country] = []
+      if not port in top_ports[proto]:
+        top_ports[proto][port] = 0
+      top_ports[proto][port] += 1
+      logs_per_country[proto][country].append((timestamp, port))
+  return logs_per_country
+
+whois_db = pickle.load(open(WHOIS_FILE, 'rb'))
[…]

### 2012-01-23 Paper_honeypot.tex
Tocheck verwijderd

@@
 \section{Introduction}
-\tocheck{Kijken of the secties nog kloppen!}
 In this paper we will try to present an overview of the data we have collected from the honeypot and explain the different attacks that were waged against it. 
